### 2018-08-01 paper/abstract.tex
initial files

In this paper, new contributions to testing Communicating Sequential
Processes (CSP) are presented, with focus on the generation of complete,
finite test suites. A test suite is complete if it can uncover every
conformance violation of the system under test with respect to a reference
model. Both reference models and implementation behaviours are represented as
CSP processes. As conformance relation, we consider trace equivalence and
trace refinement, as well as failures equivalence and failures refinement.
Complete black-box test suites here rely on the fact that the SUT's true
behaviour is represented by a member of a fault-domain, that is, a collection
of CSP processes that may or may not conform to the reference model. We
define fault domains by bounding the number of excessive states occurring in
a fault domain member's representation as a normalised transition graph, when
comparing it to the number of states present in the graph of the reference
model. This notion of fault domains is quite close to the way they are
defined for finite state machines, and these fault domains guarantee the
existence of {\it finite} complete test suites.