### 2018-08-27 paper/abstract.tex
Abstract updated to a shorter version

In this paper, two new contributions for model-based testing  
with   Communicating Sequential Processes (CSP) are presented. 
For a given CSP process representing the reference model, test suites 
checking the conformance relations 
trace refinement and  failures refinement are constructed, and their finiteness
and completeness (i.e. capability to uncover conformity violations) is proven. While complete test suites for CSP processes have been previously investigated by several authors, a sufficient condition for their finiteness is presented here for the first time. Moreover, an optimal technique for
testing the admissibility of an implementation's nondeterministic behaviour is described.