### 2018-08-27 paper/introduction.tex
updates

% =========================================================================

\section{Introduction}
\label{sec:intro}

% =================================================================================

\subsubsection*{Motivation}

Model-based testing (MBT) is an  active research field that is currently
evaluated and integrated into industrial verification processes by many
companies worldwide. This holds particularly for the embedded and
cyber-physical systems domains, where critical systems require rigorous
testing.

In the safety-critical domains, test suites with guaranteed fault coverage
are of particular interest. For black-box testing, guarantees can be given
only if certain hypotheses are satisfied. These hypotheses are usually
specified by a \emph{fault domain}:~a set of models that may or may not
conform to a given reference model. 
The so-called \emph{complete} test strategies guarantee
to accept every  system under test (SUT) conforming to the reference model, while  uncovering every conformance violation, provided that the true SUT behaviour is captured by a member of the
fault domain.

Generation methods for complete test suites have been developed for various
modelling formalisms. In this paper, we use \emph{Communicating Sequential
Processes (CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature
process-algebraic approach that has been shown to be well-suited for the
description of reactive control systems in many publications over almost five
decades. Many of these applications have
been described in~\cite{Roscoe2010} and in the references there.
Industrial success has also been reported, see, for example, \cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.

% ==================================================================================

\subsubsection*{Contributions}

This paper presents complete black-box test suites for software and systems
modelled using CSP.
They can be generated for divergence-free\footnote{The assumption of divergence freedom is usually
applied in black-box testing, since it cannot distinguish between divergence
and deadlock.} finite-state CSP processes with finite alphabets,   
interpreted both in the trace and the failures semantics.
Our results complement work published in~\cite{DBLP:conf/pts/CavalcantiS17}.
There, fault domains are specified as collections of processes refining a
``most general'' fault domain member. With that concept, complete test suites
may be finite or infinite. This gives important insight into the theory of
fault-domain testing for CSP, but we are particularly interested in {\it
finite} suites when it comes to practical application.
While~\cite{DBLP:conf/pts/CavalcantiS17} may require additional criteria to
select tests from still infinite test suites, here, we further restrict fault
domains using a graph representation of processes (originally elaborated
for the purpose of model checking) to
obtain test suites that are finite.

Our complementary approach to the definition of CSP fault domains is
presented in this paper. We observe that the failures semantics of 
every finite-state CSP process can
be represented as a finite normalised transition graph, whose edges are
labelled by the events the process engages in, and whose nodes are labelled
by minimal acceptances or, alternatively, maximal
refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals express the
degree of nondeterminism present in a given process state that is in
one-one-correspondence to a node of the normalised transition graph. Inspired
by the way that fault-domains are specified for finite state machines (FSMs),
we define them here as the set of CSP processes whose normalised transition graphs
do not exceed the order (i.e.~the number of nodes) 
of the reference model's graph by more than a given
constant.

The main contribution of this paper consists in the construction of two test suites
verifying the conformance relations \emph{trace refinement} and  \emph{failures refinement}, respectively, for any given non-divergent, finite-state, finite alphabet 
CSP reference process. We prove their completeness with respect to fault domains
of the described type.  The
existence of -- possibly infinite -- complete test suites has been
established for process algebras, and for CSP in particular, by several
authors~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/pts/CavalcantiS17}.
To the best of our knowledge, however, this article is the first to present
{\it finite}, complete test suites associated with this class of fault
domains and conformance relations.

Moreover, our result is not a simple transcription of existing knowledge
about finite, complete test suites for finite state machines: the
capabilities of CSP to express nondeterminism in a more fine-grained way than
it is possible for FSMs requires a more complex approach to testing systems
for conformity in the failures-model than it is required for model-based
testing against nondeterministic FSMs, as published, for example,
in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. We show that the approach to testing the admissibility of an SUT's nondeterministic behaviour is optimal in the
sense that it cannot be established with fewer test cases.

%It should be noted that the presented results are of a theoretical
%nature:~despite being finite, the resulting test suite size will still be too
%large to be applied in practice to real-world problems. We discuss a number
%of promising options how the test suite size can be reduced to become
%practically applicable.

% ==================================================================================

\subsubsection*{Overview}
In Section~\ref{section:preliminaries},
we present the background material relevant to our work. In Section~\ref{sec:finitecompletefails}, finite complete test suites for verifying failures refinement are presented. Test suites checking trace refinement may be regarded as a simplified version of the former class, they are presented in Section~\ref{sec:finitecomplete}. A sample test suite is presented in Section~\ref{sec:case}. The results obtained are discussed in Section~\ref{sec:conc}.

References to related work are given throughout the paper where appropriate.

% ==================================================================================