### 2018-08-28 paper/preliminaries.tex

% =========================================================================
\section{Preliminaries}
\label{section:preliminaries}
% =========================================================================

% =========================================================================
\subsection{CSP and Refinement}

%First, we introduce the CSP notation and refinement notions. We then discuss
%a graph-based semantics for finite-state CSP processes.

% =========================================================================
%\subsubsection*{CSP}
%\label{section:CSP}

\subsubsection*{Communicating Sequential Processes (CSP)} is a process algebra 
supporting system development by refinement. Using CSP, we model both systems and
their components using processes. They are characterised by their patterns
of interactions, modelled by synchronous, instantaneous, and atomic events.

A prefixing operator $e \then P$ defines a process that is ready to engage in
the event $e$, pending agreement of its environment to synchronising. After
$e$ occurs, the process behaves as defined by $P$. The environment can be
other processes, in parallel, or the environment of a system as a whole.

Two forms of choice support branching behaviour. An external choice $P
\extchoice Q$ between processes $P$ and $Q$ offers to the environment the
initial events of $P$ and $Q$. Once synchronisation takes place, the process
that offered this event is chosen and the other is discarded. In an internal
choice $P \intchoice Q$, the environment does not have an opportunity to
interfere:~the choice is made by the process itself.

\begin{example}\label{example:CSP}
  We consider the processes $P$, $Q$, and $R$ below. $P$ is initially
  ready to engage in the event $a$, and then makes an internal  choice to
  behave like either $Q$ or $R$.
  \begin{eqnarray*}
  P & = & a \then (Q\intchoice R)
  \\
  Q & = & a \then P \extchoice c \then P
  \\
  R & = & b \then P \extchoice c \then R
  \end{eqnarray*}
  $Q$, for instance, offers to the environment the choice to engage in $a$
  again or $c$. In both cases, afterwards, we have a recursion back to $P$.
  In $R$, if $b$ is chosen, we also have a recursion back to $P$. If $c$ is
  chosen, the recursion is to $R$.
  \xbox
\end{example}
%
Iterated forms $\Extchoice i: I @ P(i)$ and $\Intchoice i: I @ P(i)$ of the
external and internal choice operators allow us to define a choice over a
collection of processes $P(i)$. By convention, these choices evaluate to the process $\Stop$ 
which does not engage into any event, if the index set $I$ is empty.

There are several parallelism operators. A widely used form of parallelism $P
\parallel[cs] Q$ defines a process in which the behaviour is characterised by
those of $P$ and $Q$ in parallel, synchronising on the events in the set
$cs$. Other forms of parallelism can be defined using this operator.

Interactions that are not supposed to be visible to the environment can be
hidden. The operator $P \hide H$ defines a process that behaves as $P$, with the
interactions modelled by events in the set $H$ hidden. Frequently, hiding is used
in conjunction with parallelism: it may be desirable to make interal actions of each process in a network invisible, while events happening at their interfaces remain observable.

%Frequently, hiding is
%used in conjunction with parallelism, so that communications in a network of
%parallel processes used just to coordinate them are affected
%by the environment.

A rich collection of process operators allows us to define networks of
parallel processes in a concise and elegant way, and reason about safety,
liveness, and divergences.  A comprehensive account of the notation is given
in~\cite{Roscoe2010}.

A distinctive feature of CSP is its treatment of refinement~(as opposed to
bisimulation), which is convenient for reasoning about program correctness,
due to its treatment of nondeterminism and divergence.  A variety of semantic
models capture different notions of refinement. The simplest model
characterises a process by its possible \emph{traces}; the set $\trc(P)$ denotes
 the sequences of (non-hidden) 
events in which
$P$ can engage.  We say that a process \emph{$P$ is trace-refined by another process
$Q$}, written $P \lessdet_T Q$, if $\trc(Q)\subseteq \trc(P)$.

In fact, in every semantic model, subset containment is used to define refinement. The
model we focus on first is the failures model, which captures both
sequences of interactions and deadlock behaviour. A \emph{failure} of a process $P$
is a pair $(s,X)$ containing a trace $s$ of $P$ and a \emph{refusal}:~a set $X$ of
events in which $P$ may refuse to engage, after having performed the events of
$s$. The failures model of a process $P$ records all its failures in a set
$\fails(P)$. The semantic rules specify for each operator, how the failures of 
the resulting process can be calculated from the failures of each operand. For example, 
$\fails(P\intchoice Q) = \fails(P)\cup\fails(Q)$; see~\cite[p.~210]{Roscoe:1997:TPC:550448} for the comprehensive list of these rules.

Using notation $P/s$ to denote the process $P$ after having engaged into trace $s$,
the set 
$\refs(P/s) \defs \{~X | (s,X) \in \fails(P)~\}$ contains the  refusals
of $P$ after the
trace $s$. 
Refusals are subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}: if $(s,X)$ is a failure of $P$ and $Y\subseteq X$,
then $(s,Y)\in\fails(P)$ and $Y\in\refs(P/s)$ follows.

Failures refinement, $P \lessdet_F Q$, is defined by set containment 
$\fails(Q)\subseteq\fails(P)$. Since refusals are subset-closed, this implies
$(s,\varnothing)\in\fails(P)$ for all traces $s\in\trc(Q)$. As a consequence, failures 
refinement implies trace refinement. Therefore, using the auxiliary 
conformance relation
\begin{equation}\label{eq:conf}
  Q\ conf\ P \defs \forall s\in \trc(P) \cap \trc(Q): \refs(Q/s)
  \subseteq \refs(P/s),
\end{equation}
failures refinement can be expressed by $\lessdet_T$ and $conf$.
\begin{equation}\label{thm:fref-tref-conf}
(P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)
\end{equation}

Throughout this paper, the alphabet of the processes, that is the set of
events that are in scope, is denoted by $\Sigma$ and supposed to be finite. 
%Since we are only
%considering ``classical'' CSP processes, the alphabet is always finite. 
The
FDR tool~\cite{fdr} supports model checking and semantic analyses of
finite-state CSP processes.

For finite CSP processes, since the refusals of each process state are
subset-closed, $\refs(P/s)$ can be
constructed from the set of \emph{maximal refusals}; these are defined by 
%
\begin{equation}
\maxrefs(P/s) = \{ R \in\refs(P/s)~|~\forall R'\in \refs(P/s) - \{ R\}: R \not\subseteq R'\}
\end{equation}
%
Conversely, with the maximal refusals $\maxrefs(P/s)$ at hand, we can
reconstruct the refusals $\refs(P/s)$ by  
%
\begin{equation}
\refs(P/s) = \{ R'\in\power(\Sigma)~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
\end{equation}
%
Deterministic process states $P/s$ have exactly the one maximal refusal
$\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of $P/s$, i.e.,   the events that $P/s$ may engage into.
The maximal refusals in combination with the initials of a process express its degree of 
nondeterminism.
\begin{example}
\label{ex:nondetdegree}
Process $P = (\Stop \intchoice Q)$ has maximal refusals 
$\maxrefs(P) = \{ \Sigma \}$, because $\Stop$ 
refuses to engage into any event, and this carries over to $P$, when applying the internal 
choice operator. However, $P$ is distinguished from $\Stop$ by its initials: 
$[P]^0 = [ \Stop \intchoice Q]^0 = [Q]^0$, so $P$ may engage nondeterministically in
any initial event of $Q$, but can also refuse everything, due to internal selection of the $\Stop$ process.

Assuming that $\Sigma = \{a,b,c,d\}$, process
$Q = (e:\{a,b\} \then \Stop) \intchoice (e:\{c,d\} \then \Stop)$ has maximal refusals
$\maxrefs(Q) = \{ \{c,d\}, \{a,b\} \}$ and initials $[Q]^0=\Sigma$, so, in contrast to
$P$, nondeterminism is reflected
here by the presence of two maximal refusals.
\xbox
\end{example}

% =========================================================================
\subsubsection*{Normalised Transition Graphs for CSP Processes}
\label{sec:ntg}

As shown in~\cite{Roscoe:1994:CME:197600}, the failures semantics of 
any finite-state CSP process $P$
can be represented by a \emph{normalised transition graph}
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
partial \emph{transition function} $t$ maps a node $n$ and an event
$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
$t$, then there is a transition, that is, an outgoing edge, from $n$ with label $e$, leading to node $t(n,e)$. Normalisation of
$G(P)$ is reflected by the fact that $t$ is a function. The graph construction
in~\cite{Roscoe:1994:CME:197600} implies that all nodes $n$ in $N$ are reachable by
sequences of edges labelled by $e_1\dots e_k$ and connecting states
$\ii n,n_1,\dots,n_{k-1},n$, such that
\[
n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
n= t(n_{k-1},e_k).
\]
%
By construction, $s\in\Sigma^*$ is a trace of $P$, if and only if 
there is a path through $G(P)$ starting  at $\ii n$ whose edge labels
coincide with $s$. In analogy to $\trc(P)$, we use
notation $\trc(G(P))$ for the set of finite, initialised paths
through $G(P)$, each path represented by its finite sequence of edge labels, and we
note that $\trc(P) = \trc(G(P))$. 
Since $G(P)$ is normalised, there is a unique node
reached by applying the events from $s$ one by one, starting in $\ii n$.
Therefore, $G(P)/s$  is also well defined. By $[n]^0$ we denote the
\emph{initials} of $n$:~the set of events occurring as labels in any outgoing
transitions.
$$
[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
$$
The graph construction from~\cite{Roscoe:1994:CME:197600} guarantees that
$[G(P)/s]^0 = [P/s]^0$ for all traces of $P$.

The total function $r$ maps each node $n$ to a non-empty set of  (possibly empty) subsets of $\Sigma$. 
The graph construction guarantees that $r(G(P)/s)$ represents the maximal refusals
of $P/s$ for all 
$s\in\trc(P)$. As a consequence,  
\begin{equation}\label{eq:refsAndR}
(s,X)\in\fails(P) \iff s\in\trc(G(P)) \wedge \exists R\in r(G(P)/s): X\subseteq R,
\end{equation}
so $G(P)$ allows us to re-construct the failures semantics of $P$.

%
%
%To see that this approach works only for finite CSP processes, we consider
%the example where $\Sigma$ is infinite. In this case,
%$\maxrefs(\Stop/\langle\rangle)$ is empty, and so we cannot use this set to
%calculate the refusals of $\Stop$, that is, $\refs(\Stop/\langle\rangle)$ as
%defined above. As with refusals, we also use the transition graph-oriented
%notation $\maxrefs(n) \subseteq r(n)$ to denote the maximal refusals
%associated with graph state $n$: if $n$ is the state reached in the
%transition graph by following the edge labels in trace $s$, then $\maxrefs(n)
%= \maxrefs(P/s)$.

%Well-formed normalised transition graphs must not refuse an event of the
%initials of a state in {\it every} refusal applicable in this state; more
%formally,
%%
%\begin{equation}
%\label{eq:wellformedg}
%\forall n\in N, e\in\Sigma: (n,e)\in\dom~t \Rightarrow
%\exists R\in r(n): e\not\in R
%\end{equation}
%%
%
%
%By construction, normalised transition graphs reflect the \emph{failures
%semantics} of finite-state CSP processes:~the traces $s$ of a process are
%defined by the sequences of transitions associated with paths through its
%graph, starting at $\ii n$. The pairs $(s,R)$ with $s\in\trc(P)$ and $R\in
%r(G(P)/s)$ represent the failures of $P$.

% =========================================================================
\subsubsection*{Acceptances}
\label{sec:accs}

When investigating  tests for failures refinement, the notion of
\emph{acceptances}~\cite{Hennessy:1988:ATP:50497}, which is dual to refusals,
is also useful. An acceptance set of $P/s$ is a subset of its initials
$[P/s]^0$; equivalently, it is a subset of events labelling  outgoing transitions of
$G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance
equals $[P/s]^0$, because $P/s$ never refuses any of the events contained in
this set. If $P/s$ is nondeterministic, it internally chooses one of its
\emph{minimal acceptance sets} $A$ and never refuses any event in $A$, while
{\it possibly} refusing the events from $[P/s]^0 - A$ and 
{\it always} refusing those from $\Sigma - [P/s]^0$.  
The acceptances of $P/s$ are denoted by
$\accs(P/s)$, and the minimal acceptances by $\minaccs(P/s)$. They satisfy
the following properties.
%
\begin{eqnarray}
A\in \minaccs(P/s) & \Leftrightarrow & \exists R\in\maxrefs(P/s) \wedge A = \Sigma-R
\label{eq:accref}
\\
\bigcup \{ A~|~A\in \accs(P/s)\} & = & [P/s]^0
\label{eq:accinitials}
\\
 X\in\accs(P/s) &\Leftrightarrow & A\in \minaccs(P/s) \wedge A\subseteq X \subseteq [P/s]^0
 \label{eq:accsubset}
\end{eqnarray}
%
%Note that (\ref{eq:accref}) can be regarded as a definition of minimal
%acceptances by means of maximal refusals. In this case,
%(\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences of this
%definition and the fact that refusals are subset-closed. \fixme{alcc: are you
%sure? (4) does not talk about Acc. Is this important?}
%\fixme{jp: I'm sure and I can prove this. However, we don't need this fact, so I suggest to
%removed it.}
%
Exploiting formulas (\ref{eq:accref}), (\ref{eq:accinitials}), and
(\ref{eq:accsubset}), every node of a normalised transition graph can
alternatively be labelled with their minimal acceptances, and this
information is equivalent to that contained in the maximal refusals. Since
process states $P/s$ are equivalently expressed by states $G(P)/s$ of $P$'s
normalised transition graph, we also write $\minaccs(G(P)/s)$ and
note that (\ref{eq:refsAndR}) and (\ref{eq:accref}) imply 
\begin{equation}\label{eq:maxrefsminaccs}
\minaccs(G(P)/s) = \{ \Sigma - R | R\in r(G(P)/s) \}.
\end{equation}

% .....................................................................................
 \begin{figure}[t]
   %%\hspace*{-40mm}
   \begin{center}
     \includegraphics[width=\textwidth]{q0.pdf}
   \end{center}
   %%\vspace*{-10mm}
   \caption{Normalised transition graph of CSP process $P$ from Example~\ref{ex:a}.}
   \label{fig:tga}
 \end{figure}
% .......................................................................................

\begin{example}\label{ex:a}
Consider the CSP process $P$ in Example~\ref{example:CSP}. Its transition
graph $G(P)$ is shown in Fig.~\ref{fig:tga}. Process state $P/\varepsilon$ 
($\varepsilon$ denotes the empty trace)
is represented there as Node\_0, with $\{ a\}$ as the only minimal
acceptance, since $a$ can never be refused, and no other events are accepted.
Having engaged into $a$, the transition emanating from Node\_0 leads to
Node\_2 representing  the process state $P/a = Q\intchoice R$. The internal
choice operator induces several minimal acceptances derived from $Q$ and $R$.
Since these processes accept their initial events in external choice,
$Q\intchoice R$ induces minimal acceptance sets $\{a,c\}$ and $\{b,c\}$. Note
that event $c$ can never be refused, since it is contained in each minimal
acceptance set.

Having engaged into $c$, the next process state is represented by Node\_1.
Due to normalisation, there is only a single transition satisfying
$t(\text{Node\_2},c) = \text{Node\_1}$. This transition, however, can have
been caused by either $Q$ or $R$ engaging into $c$, so Node\_1 corresponds to
process state $Q/c \intchoice R/c = P \intchoice R$. This is reflected by the
two minimal acceptance sets labelling Node\_1.
Similar considerations explain the other nodes and transitions in
Fig.~\ref{fig:tga}.

Note that the node names including their number suffixes are generated by the
FDR tool. The numbering is generated during the normalisation procedure. So,
the node numbers do not reflect the distance from the initial node Node\_0.
\xbox
\end{example}

Summarising, refinement relations between finite-state CSP processes $P, Q$ can be be
expressed by means of their normalised transition graphs 
$$G(P) = ( N_P, {\ii n}_P, \Sigma, t_P : N_P\times\Sigma \pfun N_P, r_P : N_P \fun \mathbb{P}\mathbb{P}(\Sigma))$$ 
and 
$$G(Q) = ( N_Q, {\ii n}_Q, \Sigma, t_Q : N_Q\times\Sigma \pfun N_Q, r_Q : N_Q \fun \mathbb{P}\mathbb{P}(\Sigma))$$
in the following
way. 
\begin{lemma}
  \label{lemma:tgtrcref}
  \begin{eqnarray}
  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
  \label{eq:trcrefa}
  \\
  \label{eq:failconf}
  P \lessdet_F Q & \Leftrightarrow & P \lessdet_T Q \wedge P\ conf\ Q
  \\
  \label{eq:failrefa}
  P\ conf\ Q & \Leftrightarrow & \nonumber
  \forall s\in\trc(G(Q))\cap \trc(G(P)), R_Q\in r_Q(G(Q)/s):  \nonumber
  \\ & & \tabd
  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
  \\
  \label{eq:failrefb}
   & \Leftrightarrow & 
    \forall s\in\trc(G(Q))\cap \trc(G(P)), A_Q\in\minaccs(G(Q)/s):  \nonumber
   \\ & & \tabd
  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
 \end{eqnarray}
  \xbox
\end{lemma}

Formula (\ref{eq:trcrefa}) reflects trace refinement in terms of graph traces.  Formula~(\ref{eq:failrefa}) states how the $conf$ conformance relation can be expressed
by means of the maximal refusal functions of the graphs involved, and (\ref{eq:failrefb})
states the same in terms of the minimal acceptances that can be derived from the 
maximal refusal functions by means of (\ref{eq:maxrefsminaccs}).

% =========================================================================
\subsubsection*{Reachability Under Sets of Traces}
\label{sec:V}
Given a finite-state CSP process $P$ and its normalised transition graph
\[
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
\]
suppose that $V\subseteq\Sigma^*$ is a
prefix-closed set  of sequences of events. By $t(\ii n,V)$ we denote the set
\[
t(\ii n,V) = \{ n\in N~|~\exists s\in V: s\in\trc(P)\wedge G(P)/s = n \}
\]
of nodes in $N$ that are reachable in $G(P)$ by applying traces of $V$. The
lemma below specifies a construction method for such sets $V$ reaching {\it
every} node of $N$.

\begin{lemma}
\label{lemma:extendV} Let $P$ be a CSP process with normalised transition
graph $G(P)$. %\fixme{alcc: can a
%normalised graph for a process have states that are unreachable?}
%\fxnote{jp: You are rights: all states in G(P) are reachable by construction of the graph -- I have added this information to the section introducing the graphs.}
Let
$V\subseteq\Sigma^*$ be a finite prefix-closed set of sequences of events.
Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$, that is, $|t(\ii
n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences from $V$, extended
by any event of $\Sigma$. Then $G(P)$ reaches at least $(k+1)$ nodes under
$V\cup V.\Sigma$.
\end{lemma}
\begin{proof}
Suppose that $n'\in (N - t(\ii n,V))$.  Since all nodes in $N$ are reachable,
there exists a trace $s$ such that $G(P)/s = n'$. Decompose $s = s_1.e.s_2$
with $s_i\in\Sigma^*, e\in\Sigma$, such that $G(P)/s_1 \in t(\ii n,V)$ and
$G(P)/s_1.e \not\in t(\ii n,V)$. Such a decomposition always exists, because
$V$ is prefix-closed and therefore contains the empty trace $\varepsilon$.
Note, however, that it is not necessarily the case that $s_1\in V$.

Since $G(P)$ reaches $G(P)/s_1$ under $V$, there exists a trace $u\in V$ such
that $G(P)/u = G(P)/s_1 = \ol n$. Since $s = s_1.e.s_2$ is a trace of $P$ and
$G(P)/s_1 = \ol n$, then $(\ol n,e)$ is in the domain of $t$. So, $ G(P)/u.e
= G(P)/s_1.e = n$ is a well-defined node of $N$ not contained in $t(\ii
n,V)$. Since $u.e\in V\cup V.\Sigma$, $G(P)$ reaches at least the additional
node $n$ under $V\cup V.\Sigma$. This completes the proof. \xbox
\end{proof}

% =========================================================================
\subsubsection*{Graph Products}
\label{sec:GP}

For proving our main theorems, it is necessary to consider the \emph{product}
of normalised transition graphs. We need this only for the investigation of
corresponding traces in reference processes and processes for SUTs.
Therefore, the labelling of nodes with maximal refusals or minimal
acceptances are disregarded in the product construction. Consider two
normalised transition graphs
\[
G_i = ( N_i, \ii n_i, \Sigma, t_i : N_i\times\Sigma \pfun N_i, r_i : N_i \fun \mathbb{P}\mathbb{P}(\Sigma)),\qquad i = 1,2,
\]
over the same alphabet $\Sigma$. Their product is defined by
\begin{eqnarray}
G_1\times G_2 & = & (N_1\times N_2,(\ii n_1,\ii n_2), t:(N_1\times N_2)\times\Sigma\pfun (N_1\times N_2))
\\
\dom~t & = & \{ ((n_1,n_2),e)\in (N_1\times N_2)\times\Sigma~|   \nonumber
\\ & & \tabc
(n_1,e)\in\dom~t_1\wedge
(n_2,e) \in\dom~t_2    \}
\\
t((n_1,n_2),e) & = & (t_1(n_1,e),t_2(n_2,e))\ \text{for $((n_1,n_2),e)\in\dom~t$}
\end{eqnarray}

The following lemma will be used in the proof of our main theorem.
\begin{lemma}\label{lemma:reachproduc}
If $G_1$ has $p$ states and $G_2$ has $q$ states, then every reachable state
$(n_1,n_2)$ of the product graph $G_1\times G_2$ can be reached by a trace
%%%$s\in\trc(G_1)\cap \trc(G_2)$  we don't need this
of maximal length $(pq-1)$.
\end{lemma}
\begin{proof}
The product graph $G_1\times G_2$ has at most $pq$ states. The empty trace $\varepsilon$
reaches its initial state $(\ii n_1,\ii n_2)$. Applying Lemma~\ref{lemma:extendV}
$(pq-1)$ times with $V=\{\varepsilon \}$ implies that $G_1\times G_2$ reaches
all of its reachable states (there are at most $pq$ of them) under
$V' = V \cup V.\Sigma\cup\dots \cup V.\Sigma^{(pq-1)}$. The maximal length of traces in
$V'$ is $(pq-1)$.
\xbox
\end{proof}

% =========================================================================
\subsection{Minimal Hitting Sets}
\label{sec:hit}
% =========================================================================

The main idea of the underlying test strategy for failures refinement can be
based on solving a \emph{hitting set problem}. Given a finite collection of
finite sets $C = \{ A_1,\dots,A_n\}$, such that each $A_i$ is a subset of a
universe $\Sigma$, a \emph{hitting set} $H\subseteq\Sigma$ is a set
satisfying the following property.
%
\begin{equation}
  \label{eq:hit}
  \forall A\in C: H\cap A \neq\varnothing.
\end{equation}
%
A \emph{minimal hitting set} is a hitting set that cannot be further reduced
without losing the characteristic property (\ref{eq:hit}). By $\minhits(C)$ we denote the collection of minimal hitting sets for $C$. For the pathological case where $C$ contains an empty set, $\minhits(C)$ is also empty.
 The problem of
determining minimal hitting sets is known to be %%% \cite{Book1975-BOOKRM,5533149}
NP-hard~\cite{5533149}, but we will see below that it reduces the
effort of testing for failures refinement from a factor of $2^{|\Sigma|}$ to a
factor that equals the number of minimal hitting sets.

The following lemma establishes the fact that the $conf$ relation specified in (\ref{eq:conf}) can be chracterised by means of minimal acceptances and their
minimal hitting sets.
\begin{lemma}
\label{lemma:hseta}
Let $P, Q$ be two finite-state CSP processes.
%% satisfying $P\lessdet_T Q$.
For each $s\in\trc(P)$,
let $\minhits(P/s)$ denote the
collection of all minimal hitting sets of $\minaccs(P/s)$.
Then the following statements are equivalent.
\begin{enumerate}
\item $P\ conf\ Q$

\item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \minhits(P/s)$, $H$ is
a (not necessariliy minimal) hitting set of $\minaccs(Q/s)$.
\end{enumerate}
\end{lemma}
\begin{proof}
For showing ``$1 \Rightarrow 2$'', assume   $P\ conf\ Q$ and  
  $s\in\trc(P)\cap \trc(Q)$. Lemma~\ref{lemma:tgtrcref},
(\ref{eq:failrefb}), states that
\[
\forall A_Q\in\minaccs(G(Q)/s):
\exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
\]
Therefore, $H \in  \minhits(P/s)$ not only implies $H\cap
A_P\neq\varnothing$ for all minimal acceptances $A_P$, but also $H\cap
A_Q\neq\varnothing$ for every minimal acceptance $A_Q$, because $A_P\subseteq
A_Q$ for at least one $A_P$. As a consequence, each $H \in
\minhits(P/s)$ is also a hitting set for $\minaccs(G(Q)/s)$.

To prove ``$2 \Rightarrow 1$'', assume that that Statement~2 of the lemma holds but 
that  $P\ conf\ Q$ does {\it not} hold. 
According to Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb}), there
exists $s\in\trc(P)\cap \trc(Q)$ such that
\[
\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s): A_P\not\subseteq A_Q
\qquad (*)
\]
Let $A$ be such an acceptance set $A_Q$ fulfilling (*).
Define
\[
\overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
\]
Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),
$\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
intersection with $A$. 
Minimising $\overline H$ yields   a minimal hitting set $H\in
\minhits(P/s)$ which is {\it not} a hitting set of $\minaccs(G(Q)/s)$, a
contradiction to Assumption~2. This completes the proof of the lemma. \xbox
\end{proof}
%

It is important to point out that $\minaccs(P) = \{ \varnothing \}$ if $P = Q\intchoice \Stop$. Since $\Stop$ accepts nothing, its minimal acceptance is $\varnothing$, and this carries over to $Q\intchoice \Stop$.  From (\ref{eq:failrefb}) we conclude that 
$\varnothing\in\minaccs(P)$ implies $\minaccs(P) = \{ \varnothing \}$. These considerations
clarify that $\minhits(P/s)$ is empty if and only if  $\minaccs(P) = \{ \varnothing \}$.
We note that the proof of Lemma~\ref{lemma:hseta} also covers the situations where
$\minaccs(P/s) = \{ \varnothing \}$ and therefore  $\minhits(P/s) = \varnothing$.