### 2018-08-29 paper/conclusion.tex
Completed the discussion - only the discussion part about fault domains is still missing.

\section{Discussion}
\label{sec:conc}
% ==============================================================================
\subsubsection*{Discussion of Fault Domains}

@todo

% ==============================================================================
\subsubsection*{Discussion of Further Reductions of the Test Effort}
It is known from complete testing strategies for finite state machines that 
the upper bound $pq$ for the lengths of the 
traces used in our tests to investigate the SUT behaviour can be reduced. It is also
known from FSM testing that it is not necessary to test {\it all} traces up to this maximal length. Notable complete 
strategies supporting this fact have been presented, for example,
 in~\cite{hierons_testing_2004,DBLP:conf/forte/DorofeevaEY05,petrenko_testing_2011,simao_reducing_2012}. From~\cite{Huang2017} it
 is known that complete 
 FSM testing theories can be translated to other formalisms, such as 
Extended Finite State Machines, Kripke Structures, or CSP, resulting in likewise complete
test strategies for the latter. We intend to study translations of several promising FSM
strategies to CSP in the future. These will effectively reduce the upper bound 
$\ell\le |\Sigma|^k$ introduced in Section~\ref{sec:complexity}. The bound $h$ for
the number of sets to be used in probing the SUT for illegal deadlocks, however, cannot be further reduced, as has been established in Lemma~\ref{lemma:hseta}.

% ==============================================================================
\subsubsection*{Discussion of Adaptive Test Cases}
The tests suggested 
in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} were \emph{preset} in the
sense that the trace to be executed was pre-defined for each test. As a consequence,
the authors of~\cite{DBLP:conf/icfem/CavalcantiG07} introduced \emph{inconclusive}
as a third test result, applicable to the situations where the intended trace
of the execution was blocked, due to legal, but nondeterministic behaviour of the
SUT. In contrast to that, 
our test cases
specified in Section~\ref{sec:finitecompletefails} and \ref{sec:finitecomplete} are
adaptive. This has the advantage that test executions $(Q\parallel[\Sigma] U_F(p))$
for failures refinement
will never block before the final step specified by branch (\ref{eq:ufd}). As
a consequence, we do not need inconclusive test results. It should be noted, however,
that it is necessary for our test verdicts to recognise also deadlocks in the final
test step and interpret 
them as FAIL, as described in Section~\ref{sec:finitecompletefails}. In practice,
this is realised by adding a timeout event to the testing environment which indicates
deadlock situations. For real-time systems, this is an accepted technique, because
the SUT has to respond within a pre-defined latency interval, otherwise its behaviour 
is considered to be blocked and regarded as a failure. Our tests executions 
$(Q\parallel[\Sigma] U_T(p))$ for trace refinement never block at all. The adaptive       
behaviour of our test cases, however, induce the obligation to check that {\it all}
possible executions have been performed before the test can be considered as passed.
Typically, it is therefore assumed that a 
\emph{complete testing assumption}~\cite{hierons_testing_2004} holds which means that
every possible behaviour of the SUT has been performed after a finite number of
test executions. 
In practice, this is realised by executing each test several times, 
recording the traces that have been performed, and using hardware or software coverage 
analysers to determine whether all possible test execution behaviours of the SUT
have been observed. Therefore, the application of adaptive test cases comes at the price
of having to apply some grey-box testing techniques enabling us to decide whether
all SUT behaviours have been observed.

% ==============================================================================
\subsubsection*{Implications for CSP Model Checking}
As explained in the previous sections, passing a test is specified by
$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k)) \hide \Sigma$
for failures testing. If the SUT $Q$ is not a programmed piece of software or
an integrated HW/SW system, but just another CSP process specification,
it is of course possible to verify these pass criteria using the FDR4 model checker.
For checking the refinement relation $P\lessdet_F Q$, the pass criteria have to be verified
for $k=0,\dots,pq-1$, where $p$ and $q$ indicate the number of nodes in $P$'s transition
graph and the maximal number of nodes in $Q$'s graph, respectively (Theorem~\ref{th:failurestest}). Since the test cases
$U_F(k)$ have such a simple structure, it is an interesting question for further 
research whether checking 
$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k)) \hide \Sigma$ for $k=0,\dots,pq-1$ can be faster than directly 
checking $P\lessdet_F Q$, as one would do in the usual 
approach with FDR4. This is of particular interest, since the checks could be parallelised 
on several CPUs. Alternatively it is interesting to investigate whether the
the check\footnote{We are grateful to Bill Roscoe for suggesting this option.}
\[
(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] \Intchoice_{k=0}^{pq-1} U_F(k)) \hide \Sigma
\]
may perform better than the check of $P\lessdet_F Q$, since the former allows for 
other optimisations in the model checker than the latter.

For large and complex CSP processes $Q$, it may be too time consuming to generate
its normalised transition graph, so that its number $q$ of nodes is unknown. 
In such a case the suggested options may still be used
as efficient bug finders: use  $p$ of reference process $P$ as initial $q$-value
and 
increment $q$ from there, as long as each increment reveals new errors.

% ==============================================================================
\subsubsection*{Conclusion}

In this paper, we have introduced finite complete testing strategies for 
model-based testing against CSP reference models. The strategies are applicable   
to the conformance relations failures refinement and trace refinement. The underlying fault domains have been defined as the sets of CSP processes whose normalised transition
graphs did not have more than a given number of additional nodes, when compared to the transition graph of the reference process. For these domains, finite complete 
test suites are available. We have shown for the strategy checking failures refinement,
 that the way of probing the SUT for illegal deadlocks used in our test cases is 
 optimal in the sense that it is not possible to guarantee exhaustiveness with fewer
 probes.

% ==============================================================================

\paragraph{Acknowledgements}
The authors would like to thank Bill Roscoe and Thomas Gibson-Robinson for their advice on using the FDR4 model checker and for very helpful discussions concerning the potential implications of this paper in the field of model-checking.