### 2018-09-02 paper/introduction.tex
small suggestions

% =========================================================================

\section{Introduction}
\label{sec:intro}

% =================================================================================

\subsubsection*{Motivation}

Model-based testing (MBT) is an  active research field that is currently
evaluated and integrated into industrial verification processes by many
companies worldwide. This holds particularly for the embedded and
cyber-physical systems domains, where critical systems require rigorous
testing.

In the safety-critical domain, test suites with guaranteed fault coverage are
of particular interest. For black-box testing, guarantees can be given only
if certain hypotheses are satisfied. These hypotheses are usually specified
by a \emph{fault domain}:~a set of models that may or may not conform to a
given reference model. \emph{Complete} test strategies guarantee to accept
every  system under test (SUT) conforming to the reference model, and uncover
every conformance violation, provided that the SUT behaviour is captured by a
member of the fault domain.

Generation methods for complete test suites have been developed for various
formalisms. In this paper, we use \emph{Communicating Sequential Processes
(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature process
algebra that has been shown to be well-suited for the description of reactive
control systems in many publications over almost five decades. Many of these
applications are described in~\cite{Roscoe2010} and in the references there.
Industrial success has also been reported, see, for example,
\cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.

% ==================================================================================

\subsubsection*{Contributions}

This paper presents complete black-box test suites for software and systems
modelled using CSP. They can be generated for divergence-free\footnote{The
assumption of divergence freedom is usually applied in black-box testing,
since it cannot distinguish between divergence and deadlock.} finite-state
CSP processes with finite alphabets, interpreted both in the trace and the
failures semantics. Our results complement work
in~\cite{DBLP:conf/pts/CavalcantiS17}. There, fault domains are specified as
collections of processes refining a ``most general'' fault domain member.
With that concept, complete test suites may be finite or infinite. This gives
important insight into the theory of fault-domain testing for CSP, but we are
particularly interested in {\it finite} suites when it comes to practical
application. While~\cite{DBLP:conf/pts/CavalcantiS17} may require additional
criteria to select tests from still infinite test suites, here, we further
restrict fault domains using a graph representation of processes (originally
elaborated for model checking) to obtain finite test suites. \fixme{alcc: I
think termination is not handled either?}

Our complementary approach to the definition of CSP fault domains is
presented in this paper. We observe that the failures semantics of
every finite-state CSP process can
be represented as a finite normalised transition graph, whose edges are
labelled by the events the process engages in, and whose nodes are labelled
by minimal acceptances or, alternatively, maximal
refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals express the
degree of nondeterminism present in a given process state that is in
one-one-correspondence to a node of the normalised transition graph. Inspired
by the way that fault-domains are specified for finite state machines (FSMs),
we define them here as the set of CSP processes whose normalised transition graphs
do not exceed the order (i.e.~the number of nodes)
of the reference model's graph by more than a given
constant.

The main contribution of this paper is the construction of two test suites to
verify the conformance relations \emph{trace refinement} and \emph{failures
refinement} for any given non-divergent, finite-state, finite alphabet CSP
process. We prove their completeness with respect to fault domains
of the described type.  The existence of -- possibly infinite -- complete
test suites has been established for process algebras, and for CSP in
particular, by several
authors~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}.
To the best of our knowledge, however, this article is the first to present
{\it finite}, complete test suites associated with this class of fault
domains and conformance relations.

Moreover, our result is not a simple transcription of existing knowledge
about finite, complete test suites for finite state machines: the
capabilities of CSP to express nondeterminism in a more fine-grained way than
it is possible for FSMs requires a more complex approach to testing systems
for conformity in the failures-model than it is required for model-based
testing against nondeterministic FSMs, as published, for example,
in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. We show that the approach to testing the admissibility of an SUT's nondeterministic behaviour is optimal in the
sense that it cannot be established with fewer test cases.

%It should be noted that the presented results are of a theoretical
%nature:~despite being finite, the resulting test suite size will still be too
%large to be applied in practice to real-world problems. We discuss a number
%of promising options how the test suite size can be reduced to become
%practically applicable.

% ==================================================================================

\subsubsection*{Overview}
In Section~\ref{section:preliminaries},
we present the background material relevant to our work. In Section~\ref{sec:finitecompletefails}, finite complete test suites for verifying failures refinement are presented. Test suites checking trace refinement may be regarded as a simplified version of the former class, they are presented in Section~\ref{sec:finitecomplete}. A sample test suite is presented in Section~\ref{sec:case}. The results obtained are discussed in Section~\ref{sec:conc}.

References to related work are given throughout the paper where appropriate.

% ==================================================================================