### 2018-09-02 paper/abstract.tex
small suggestions

In this paper, two new contributions for model-based testing with
Communicating Sequential Processes (CSP) are presented. For a given CSP
process representing the reference model, test suites checking the
conformance relations trace and failures refinement are constructed, and
their finiteness and completeness (that is, capability to uncover conformity
violations) is proven. While complete test suites for CSP processes have been
previously investigated by several authors, a sufficient condition for their
finiteness is presented here for the first time. Moreover, an optimal
technique for testing the admissibility of an implementation's
nondeterministic behaviour is described.