### 2018-09-02 paper/introduction.tex

% =========================================================================

\section{Introduction}
\label{sec:intro}

% =================================================================================

\subsubsection*{Motivation}

Model-based testing (MBT) is an  active research field that is currently
evaluated and integrated into industrial verification processes by many
companies worldwide. This holds particularly for the embedded and
cyber-physical systems domains, where critical systems require rigorous
testing.

In the safety-critical domain, test suites with guaranteed fault coverage are
of particular interest. For black-box testing, guarantees can be given only
if certain hypotheses are satisfied. These hypotheses are usually specified
by a \emph{fault domain}:~a set of models that may or may not conform to a
given reference model. \emph{Complete} test strategies guarantee to accept
every  system under test (SUT) conforming to the reference model, and uncover
every conformance violation, provided that the SUT behaviour is captured by a
member of the fault domain.

Generation techniques for complete test suites have been developed for
various formalisms. Here, we tackle \emph{Communicating Sequential
Processes~(CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature
process algebra that has been shown to be well-suited for the description of
reactive control systems in many publications over almost five decades. Many
of these applications are described in~\cite{Roscoe2010} and in the
references there. Industrial success has also been reported; see, for
example, \cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.

% ==================================================================================

\subsubsection*{Contributions}

This paper presents complete black-box test suites for software and systems
modelled using CSP. They can be generated for divergence-free\footnote{The
assumption of divergence freedom is usually applied in black-box testing,
since it cannot distinguish between divergence and deadlock.} finite-state
CSP processes with finite alphabets, interpreted both in the trace and the
failures semantics. Our results complement work
in~\cite{DBLP:conf/pts/CavalcantiS17}. There, fault domains are specified as
collections of processes refining a ``most general'' fault domain member.
With that concept, complete test suites may be finite or infinite. This gives
important insight into the theory of fault-domain testing for CSP, but we are
particularly interested in {\it finite} suites when it comes to practical
application. While~\cite{DBLP:conf/pts/CavalcantiS17} may require additional
criteria to select tests from still infinite test suites, here, we further
restrict fault domains using a graph representation of processes (originally
elaborated for model checking) to obtain finite test suites. \fixme{alcc: I
think termination is not handled either?}

Our approach to the definition of CSP fault domains is presented in this
paper. We take advantage of results on model checking of CSP
processes, where the failures semantics of a finite-state CSP process is
represented as a finite normalised transition graph, whose edges are labelled
by the events the process engages in, and whose nodes are labelled by minimal
acceptances or, alternatively, maximal
refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals express the
degree of nondeterminism present in a process state that is in
one-one-correspondence to a node of the normalised transition graph. Inspired
by the way fault-domains are specified for finite state machines (FSMs),
we define them here as the set of CSP processes whose normalised transition
graphs do not exceed the order (that is, the number of nodes) of the
reference model's graph by more than a given constant.

The main contribution of this paper is the construction of two test suites to
verify the conformance relations \emph{trace refinement} and \emph{failures
refinement} for any given non-divergent, finite-state, finite alphabet CSP
process. We prove their completeness with respect to fault domains
of the described type.  The existence of -- possibly infinite -- complete
test suites has been established for process algebras, and for CSP in
particular, by several
authors~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/icfem/CavalcantiG07,DBLP:conf/pts/CavalcantiS17}.
To the best of our knowledge, however, this article is the first to present
{\it finite}, complete test suites associated with this class of fault
domains and conformance relations.

Moreover, our result is not a simple transcription of existing knowledge
about finite, complete test suites for finite state machines: the
capabilities of CSP to express nondeterminism in a more fine-grained way than
it is possible for FSMs requires a more complex approach to testing systems
for conformity in the failures-model than it is required for model-based
testing against nondeterministic FSMs, as published, for example,
in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. We show that the
approach to testing the admissibility of an SUT's nondeterministic behaviour
is optimal in the sense that it cannot be established with fewer test cases.
\fixme{alcc: I did not understand this?}

%It should be noted that the presented results are of a theoretical
%nature:~despite being finite, the resulting test suite size will still be too
%large to be applied in practice to real-world problems. We discuss a number
%of promising options how the test suite size can be reduced to become
%practically applicable.

% ==================================================================================

\subsubsection*{Overview}
In Section~\ref{section:preliminaries}, we present the background material
relevant to our work. In Section~\ref{sec:finitecompletefails}, finite
complete test suites for verifying failures refinement are presented. Test
suites checking trace refinement are a simplified version of the former
class; they are presented in Section~\ref{sec:finitecomplete}. A sample test
suite is presented in Section~\ref{sec:case}. Our results are discussed in
Section~\ref{sec:conc}, where we also conclude. References to related work
are given throughout the paper where appropriate.

% ==================================================================================