### 2018-09-07 paper/casestudies.tex
Node number for transition graphs fixed, Example improved

% =======================================================================
\section{Testing for Failures Refinement -- an Example}
\label{sec:case}
% =======================================================================

%
%For implementing the test case $U_F(p)$ with sub-processes $U_F(p,s)$, it is
%advisable to avoid an enumeration of traces $s$ of the reference process.
%Instead, we calculate the following auxiliary functions from $P$'s transition
%graph.
%%
%\begin{eqnarray*}
%\text{\it initials} & : & N \fun \power (\Sigma)
%\\
%\minhits & : & N \fun \power\power(\Sigma)
%\end{eqnarray*}
%%
%As mentioned, in a state $n = G(P)/s$, the set $\text{\it initials}(n)$
%equals the events labelling outgoing edges of $n$, so  $\text{\it
%initials}(n) = [P/s]^0$. The function $\minhits$ maps $n$ to the set of all
%minimal hitting sets associated with the minimal acceptances of $n$, so
%$\minhits(n) = \minhits(P/s)$. Using the transition function $t$ and the
%above functions, $U_F(p)$ can be re-written as the failures-equivalent CSP
%process below.
%%
%\begin{eqnarray}
%U_F^1(p) & = & U_F^1(p,0,\ii n)
%\\
%U_F^1(p,k,n) & = & \big(\Extchoice e:(\Sigma - \text{\it initials}(n)) @ e \then \efail\then \Stop \big)
%\label{eq:xufa}
%\\ & & \extchoice \nonumber
%\\ & & (\minhits(n) = \varnothing) \& \big( \epass \then \Stop \big)
%\label{eq:xufb}
%\\ & & \extchoice \nonumber
%\\ & & (k < p) \& \big(\Extchoice e:\text{\it initials}(n) @ e \then U_F^1(p,(k+1),t(n,e)) \big)
%\label{eq:xufc}
%\\ & & \extchoice \nonumber
%\\ & & (k = p) \& \big( \sqcap_{H\in\minhits(n)} (\Extchoice e:H @ e \then \epass \then\Stop ) \big)
%\label{eq:xufd}
%\end{eqnarray}
%\fixme{alcc: I think the example should be for the theory presented before,
%not with a different notion of test. In particular, 33 is very different.}

Generating the test cases $U_F(p)$ specified in  (\ref{eq:UFP}) for the reference
process $P$ discussed in Example~\ref{example:CSP},
results in the following instantiations of initials, minimal hitting sets, and
transition function; this can be directly seen by analysing $P$'s normalised 
transition graph with nodes $N =\{0,1,2,3\}$ as shown in Fig.~\ref{fig:tga}.

\begin{eqnarray*}
{ }[0]^0 & = & \{ a \} \\
{ }[1]^0 & = & \{ a,b,c \} \\
{ }[2]^0 & = & \{ a,b,c \} \\
{ }[3]^0 & = & \{ b,c \} \\
& & \\
\minhits(0) & = & \{ \{ a\} \} \\
\minhits(1) & = & \{ \{a,b\}, \{c\}, \} \\
\minhits(2) & = & \{  \{a,b\},  \{a,c\},\} \\
\minhits(3) & = & \{ \{ b\}, \{c\} \} \\
& & \\
t(0,a) & = & 1 \\
t(1,a) & = & 0 \\
t(1,b) & = & 0 \\
t(1,c) & = & 2 \\
t(2,a) & = & 1 \\
t(2,b) & = & 0 \\
t(2,c) & = & 3 \\
t(3,b) & = & 0 \\
t(3,c) &  =& 3
\end{eqnarray*}

% .....................................................................................
 \begin{figure}
 %%\hspace*{-40mm}
 \begin{center}
\includegraphics[width=\textwidth]{z.pdf}
\end{center}
%%\vspace*{-10mm}
\caption{Normalised transition graph of faulty implementation $Z$  from Example~\ref{ex:uf1tests}.}
 \label{fig:tgZ}
 \end{figure}
% .......................................................................................

\begin{example}
\label{ex:uf1tests} Consider the following implementation $Z$ of process $P$
from Example~\ref{example:CSP} that is erroneous from the point of view of
failures refinement.
\begin{eqnarray*}
Z & = & a \then (Q_1 \intchoice R_1(r_{max},0))
\\
Q_1 & = & a\then Z \extchoice c\then Z
\\
R_1(r_{max},k) & = & (k < r_{max}) \& \big(b\then Z \extchoice  c\then R_1(r_{max},k+1)\big)
\\ & & \extchoice
\\ & & (k = r_{max}) \& \big(b\then Z \intchoice c\then R_1(r_{max},r_{max})\big)
\end{eqnarray*}
It is easy to see (and can be checked with FDR) that $Z$ is trace-equivalent
to $P$. While $k < r_{max}$, $Z$ also accepts the same sets of events as $P$.
When $R_1(r_{max},k)$ runs through several recursions and $k = r_{max}$ is
fulfilled, however, $R_1(r_{max},k)$ makes an internal choice, instead of
offering an external choice, so $P\not\lessdet_F Z$.
Fig.~\ref{fig:tgZ} shows the normalised transition graph of $Z$ for $r_{max}
= 3$.

%%%and refinement does not hold.
% alcc: I don't think there is a notion of refinement for state?
%so this process state no
%longer refines the corresponding process state $R$ of the reference process
%$P$ in the failures model.

Running the test $U_F(k)$ against $Z$ for $k=0,\dots,20$ ($G(P)$ has $p =
4$ states and $G(Z)$ has $q=5$, so $pq=20$ is an upper bound for the test
depth to be used according to Theorem~\ref{th:failurestest}), tests
$U_F(0),\dots, U_F(3)$ are passed by $Z$, but $Z$ fails $U_F(4)$,
because after  execution trace
\[
s = a.c.c.c, \qquad\text{(note that $G(P)/s = \text{node}\ 3$ according to Fig.~\ref{fig:tga})},
\]
the test $U_F(4)$ offers hitting sets from $\minhits(3) = \{\{b\},\{c\}\}$
in branch (\ref{eq:ufd}). Therefore,  there exists one test execution
where $Z/s$ accepts only $\{b\}$ due to the internal choice (note 
from Fig.~\ref{fig:tgZ} that
$G(Z)/s = 4$), while $U_F(4)/s$
only offers $\{c\}$ in branch (\ref{eq:ufd}) or $\{ a\} = \Sigma - [3]^0$ for 
branch (\ref{eq:ufa}).  As a consequence, this execution of 
$(Z\parallel[\Sigma] U_F(4))/s$ deadlocks, and the $pass$ event cannot be
produced. Another failing execution arises if $Z/s$ chooses to accept only
$\{c \}$, while $U_F(4)/s$ choses to accept only $\{a,b\}$. Therefore,
\[
(\epass\then \Stop)\not\lessdet_F  (Z\parallel[\Sigma] U_F(4)),
\]
and the test fails.
\xbox
\end{example}
%%%\fixme{alcc: I would write down the actual test generated.}

% ======================================================================