### 2018-09-30 paper/conclusion.tex
finished review

% ==============================================================================
\section{Discussion and Conclusions}
\label{sec:conc}
% ==============================================================================

% ==============================================================================
\subsubsection*{Discussion of Further Reductions of the Test Effort}
It is known from complete testing strategies for finite state machines that
the upper bound $pq$ for the lengths of the traces used in our tests to
investigate the SUT behaviour can be reduced. It is also known from FSM
testing that it is not necessary to test {\it all} traces up to this maximal
length. Notable complete strategies supporting this fact have been presented,
for example,
in~\cite{hierons_testing_2004,DBLP:conf/forte/DorofeevaEY05,petrenko_testing_2011,simao_reducing_2012}.
From~\cite{Huang2017} it is known that complete FSM testing theories can be
translated to other formalisms, such as Extended Finite State Machines,
Kripke Structures, or CSP, resulting in likewise complete test strategies for
the latter. We intend to study translations of several promising FSM
strategies to CSP in the future. These will effectively reduce the upper
bound $\ell\le |\Sigma|^k$ introduced in Section~\ref{sec:complexity}. The
bound $h$ for the number of sets to be used in probing the SUT for illegal
deadlocks, however, cannot be further reduced, as has been established in
Lemma~\ref{lemma:hseta}.

% ==============================================================================
\subsubsection*{Discussion of Adaptive Test Cases}
The tests suggested
in~\cite{Hennessy:1988:ATP:50497,DBLP:conf/icfem/CavalcantiG07} were \emph{preset} in the
sense that the trace to be executed was pre-defined for each test. As a consequence,
the authors of~\cite{DBLP:conf/icfem/CavalcantiG07} introduced \emph{inconclusive}
as a third test result, applicable to the situations where the intended trace
of the execution was blocked, due to legal, but nondeterministic behaviour of the
SUT.

In contrast to that, our test cases specified in
Section~\ref{sec:finitecompletefails} and \ref{sec:finitecomplete} are
adaptive. This has the advantage that test executions $(Q\parallel[\Sigma]
U_F(p))$ for failures refinement never blocks before the final step specified
by branch (\ref{eq:ufd}), and so we do not need inconclusive test results.
\fixme{From the proofs of the main theorems, and what you say below, I
thought you did have verdicts coming from deadlock, and you simply chose not
to mark it with an inc event.} It should be noted, however, that it is
necessary for our test verdicts to recognise also deadlocks in the final test
step and interpret them as FAIL, as described in
Section~\ref{sec:finitecompletefails}. In practice, this is realised by
adding a timeout event to the testing environment which indicates deadlock
situations. For real-time systems, this is an accepted technique, because the
SUT has to respond within a pre-defined latency interval, otherwise its
behaviour is considered to be blocked and regarded as a failure.

Our tests executions $(Q\parallel[\Sigma] U_T(p))$ for trace refinement never
block at all. The adaptive behaviour of our test cases, however, induce the
\fixme{Missing hiding. And also, if the SUT blocks, the test should block,
no?} obligation to check that {\it all} possible executions have been
performed before the test can be considered as passed. Typically, it is
therefore assumed that a \emph{complete testing
assumption}~\cite{hierons_testing_2004} \fixme{I think you need this for the
failures refinement tests as well.} holds, which means that every possible
behaviour of the SUT is performed after a finite number of test executions.
In practice, this is realised by executing each test several times, recording
the traces that have been performed, and using hardware or software coverage
analysers to determine whether all possible test execution behaviours of the
SUT have been observed. Therefore, adaptive test cases come at the price of
having to apply some grey-box testing techniques enabling us to decide
whether all SUT behaviours have been observed.

% ==============================================================================
\subsubsection*{Discussion of Fault Domains}
As already mentioned, the work in~\cite{DBLP:conf/pts/CavalcantiS17} defines
a fault domain as the set of processes that refine a given CSP process.  In
that context, only testing for traces refinement is considered, and the
complete test suites may not be finite. So, the work presented here goes well
beyond what is achieved there. On the other hand,
\cite{DBLP:conf/pts/CavalcantiS17} presents an algorithm for test generation
that can be easily adapted to consider additional selection and termination
criteria, like, for example, the length of the traces used to construct
tests. It would be possible, for instance, to use the bound indicated here.
Moreover, specifying a fault domain as a CSP process allows us to model
domain-specific knowledge using CSP. For example, if an initialisation
component defined by a process $I$ can be regarded as correct without further
testing, we can use $I; RUN$ as a fault domain, to indicate that any SUT of
interest implements $I$ correctly, but afterwards has a potentially arbitrary
behaviour specified by $RUN$. In addition, the work
in~\cite{DBLP:conf/pts/CavalcantiS17} is not restricted to finite or
non-terminating processes.

% ==============================================================================
\subsubsection*{Implications for CSP Model Checking}
As explained in the previous sections, passing a test is specified by
$(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(k)) \hide \Sigma$ for
failures testing. If the SUT $Q$ is not a programmed piece of software or an
integrated hardware or software system, but just another CSP process
specification, it is of course possible to verify the pass criterion using
the FDR4 model checker. For checking the refinement relation $P\lessdet_F Q$,
the pass criterion has to be verified for $k=0,\dots,pq-1$, where $p$ and $q$
indicate the number of nodes in $P$'s transition graph and the maximal number
of nodes in $Q$'s graph, respectively (Theorem~\ref{th:failurestest}). Since
the test cases $U_F(k)$ have such a simple structure, it is an interesting
question for further research whether checking $(\epass\then\Stop) \lessdet_F
(Q\parallel[\Sigma] U_F(k)) \hide \Sigma$ for $k=0,\dots,pq-1$ can be faster
than directly checking $P\lessdet_F Q$, as one would do in the usual approach
with FDR4. This is of particular interest, since the checks could be
parallelised on several CPUs. Alternatively it is interesting to investigate
whether the check\footnote{We are grateful to Bill Roscoe for suggesting this
option.}
\[
(\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] \Intchoice_{k=0}^{pq-1} U_F(k)) \hide \Sigma
\]
may perform better than the check of $P\lessdet_F Q$, since the former allows for
other optimisations in the model checker than the latter.

For large and complex CSP processes $Q$, it may be too time consuming to
generate its normalised transition graph, so that its number $q$ of nodes is
unknown. In such a case the suggested options may still be used as efficient
bug finders: use  $p$ of reference process $P$ as initial $q$-value and
increment $q$ from there, as long as each increment reveals new errors.

% ==============================================================================
\subsubsection*{Conclusion}

In this paper, we have introduced finite complete testing strategies for
model-based testing against finite, non-terminating CSP reference models. The
strategies are applicable to the conformance relations failures refinement
and trace refinement. The underlying fault domains have been defined as the
sets of CSP processes whose normalised transition graphs do not have more
than a given number of additional nodes, when compared to the transition
graph of the reference process. For these domains, finite complete test
suites are available. We have shown for the strategy to check failures
refinement that the way of probing the SUT for illegal deadlocks used in our
test cases is optimal in the sense that it is not possible to guarantee
exhaustiveness with fewer probes.

% ==============================================================================

\paragraph{Acknowledgements}
The authors would like to thank Bill Roscoe and Thomas Gibson-Robinson for
their advice on using the FDR4 model checker and for very helpful discussions
concerning the potential implications of this paper in the field of model
checking. The work of Cavalcanti is funded by the Royal Academy of
Engineering and UK EPSRC Grant EP/R025134/1.