### 2018-10-01 paper/abstract.tex
further points of Ana's suggestions handled

In this paper, two new contributions for model-based testing with
Communicating Sequential Processes (CSP) are presented. For a given CSP
process representing the reference model, test suites checking the
conformance relations trace refinement and failures refinement are constructed, and
their finiteness and completeness (that is, capability to uncover conformity
violations) is proven. While complete test suites for CSP processes have been
previously investigated by several authors, a sufficient condition for their
finiteness is presented here for the first time. 
%Moreover, an optimal
%technique for testing the admissibility of an implementation's
%nondeterministic behaviour is described. 
It is shown that the test suites are optimal with respect to two aspects:  (a)  the maximal length of test traces cannot be reduced, and (b) the nondeterministic behaviour cannot be tested with smaller or fewer sets of events, without losing the test suite's completeness property. \fixme{alcc: Is the technique
applying the test suites?}
\fxnote{jp: this sentence needed to be re-phrased -- the new version should be better understandable}