### 2018-10-03 paper/finitecomplete-trace.tex

% ==========================================================================
\section{Finite Complete Test Suites for CSP Trace Refinement}
\label{sec:finitecomplete}
% ==========================================================================

For establishing trace refinement, the following class of adaptive test cases
will be used for a given reference process $P$ and  integers $p \ge 0$.
Just as for the tests developed  in
Section~\ref{sec:finitecompletefails} to verify failures refinement,
the tests for trace refinement are
derived from the reference model's transition graph
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)).
$$
In contrast to the tests for failures refinement~(\ref{eq:UFP}), however, we
do not need to check the SUT with respect to its acceptance of hitting sets.
Therefore, these do not occur in the specification of the test cases below,
and we use the condition $\minaccs(n) = \{ \varnothing \}$ instead of
$\minhits(n) = \varnothing$ in branch (\ref{eq:utb}) 
to indicate that minimal hitting sets need not to
be calculated for generating these tests from $G(P)$. From
(\ref{eq:minhitminaccempty}) we know that these conditions are equivalent.

%\begin{eqnarray}
%U_T(p) & = & U_T(p,\varepsilon)
%\\
%U_T(p,s) & = & \big(\Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
%\label{eq:uta}
%\\ & & \extchoice \nonumber
%\\ & & (\minaccs(P/s) = \{ \varnothing \})   \&   \big( \epass \then \Stop \big)
%\label{eq:utb}
%\\ & & \extchoice \nonumber
%\\ & & (\#s < p) \& \big(\Extchoice e:[P/s]^0 @ e \then U_T(p,s.e) \big)
%\label{eq:utc}
%\\ & & \extchoice \nonumber
%\\ & & (\#s = p) \& \big( \epass\then \Stop  \big)
%\label{eq:utd}
%\end{eqnarray}

\begin{eqnarray}
U_T(p) & = & U_T(p,0,\ii n)
\\
U_T(p,k,n) & = & \big(  e:(\Sigma - [n]^0)   \then \efail\then \Stop \big)
\label{eq:uta}
\\ & & \extchoice \nonumber
\\ & & (\minaccs(n) = \{ \varnothing \})   \&   \big( \epass \then \Stop \big)
\label{eq:utb}
\\ & & \extchoice \nonumber
\\ & & (k < p) \& \big( e:[P/s]^0   \then U_T(p,k+1,t(n,e)) \big)
\label{eq:utc}
\\ & & \extchoice \nonumber
\\ & & (k = p) \& \big( \epass\then \Stop  \big)
\label{eq:utd}
\end{eqnarray}
%
It is easy to see that the tests $U_T(p)$ satisfy the properties
\begin{eqnarray}
\label{eq:ifpaT}
  &  & U_T(p)/s = U_T(p,\#s,G(P)/s)
\\
\label{eq:ifpbT}
e\not\in [P/s]^0 & \implies & U_T(p)/s.e = (\efail\then\Stop)
\end{eqnarray}
proven in Lemma~\ref{lemma:ufproperties} for $U_F(p)$ for traces
$s\in\trc(P)$ with $\#s \le p$. 
\fixme{In (37), I think there is an implicit assumption about
$p$ and $\# s$?}\fxnote{jp: yes, I added $\#s \le p$ in the text.}

%The difference between adaptive tests $U_T(p)$ for trace refinement and
%$U_F(p)$ for failures refinement consists in the fact that the former do not
%``probe'' the SUT with respect to minimal sets of events to be accepted
%without blocking.

% ==========================================================================
Since
$U_T(p)$ never blocks any $Q$-event before terminating, the pass criterion
can be based on trace refinement instead of failures refinement as required
in (\ref{eq:passF}).
%
\begin{equation}
\label{eq:passT}
Q\ \pass\ U_T(p) \defs (\epass\then\Stop) \lessdet_T (Q\parallel[\Sigma] U_T(p)) \hide \Sigma
\end{equation}
\fixme{The proofs don't use this definition. It should change.}
\fxnote{jp: I do not understand this remark -- we should discuss it on the phone.
Please note the addition explanations I have added here below (\ref{eq:passT}).}

If the SUT process $Q$ deadlocks internally after a trace $s$ where also the reference
process $P$ is in a state where deadlock is possible, this is captured by the fact
that $\minaccs(n) = \{ \varnothing\}$ for $n = G(P)/s$. As a consequence, branch
(\ref{eq:utb}) of a test case execution state $U_T(p,k,n)$  
with $\#s = k \le p$   can be entered
and the  test execution terminates with $\epass$. If, however,  
$Q$ blocks after a trace $s'$
where the reference process satisfies $\minaccs(P/s') \neq\varnothing$, branch 
(\ref{eq:utb})  cannot be taken, and the test execution stops without producing
$\epass$ or $\efail$. In contrast to the test for failures refinement, this is   interpreted here as a successful test execution, because unexpected blocking
of the SUT does not violate the trace refinement relation, as long as all traces 
executed by the SUT are traces of the reference process. In particular, if 
neither $\epass$ nor $\efail$ is ever produced, so that
$(Q\parallel[\Sigma] U_T(p)) \hide \Sigma = \Stop$, the test passes, because
$(\epass\then\Stop) \lessdet_T \Stop$ holds.

The existence of complete, finite test suites is expressed in analogy to
Theorem~\ref{th:failurestest}. A noteworthy difference is that the complete
suite for trace refinement just needs the single adaptive test case
$U_T(pq-1)$, while failures refinement requires the execution of $\{
U_F(0),\dots,U_F(pq-1)\}$. The reason is that $U_T(pq-1)$ identifies trace
errors for all traces up to length $pq$, while $U_F(pq-1)$ only probes for
erroneous acceptances at the end of each trace of length $(pq -1)$. 
%
\begin{theorem}\label{th:tracetest}
Let $P$ be a non-terminating, divergence-free CSP process over alphabet $\Sigma$ whose
normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
D}$ as the set of all non-terminating, divergence-free CSP processes over alphabet $\Sigma$,
whose transition graph has at most $q$ states with $q \ge p$. Then the test
suite
\[
\TS_T = \{ U_T(pq-1)   \}
\]
is complete with respect to ${\cal F} = (P,\lessdet_T,{\cal D})$.
\xbox
\end{theorem}
%\begin{proof}
%The theorem follows directly
%from Step~1 in the proof of Lemma~\ref{lemma:mainfsound} and
%Case~1 in the proof of Lemma~\ref{lemma:mainfexhaustive}.
%\xbox
%\end{proof}
%%
%Examples are provided after our discussion of size of the test suites.

As before in the context of Theorem~\ref{th:failurestest}, 
the proof of Theorem~\ref{th:tracetest} is structured into two lemmas, the first 
establishing soundness, the second exhaustiveness.

\fxnote{jp: Wen-ling and I have now provided a full proof for Theorem~\ref{th:tracetest}}
% =============================================================================== 
\begin{lemma}\label{lemma:mainfsoundtrace}
A test suite $\TS_T$ generated from a CSP process $P$, as specified in
Theorem~\ref{th:tracetest}, is passed by every CSP process $Q$ satisfying
$P\lessdet_T Q$.
\end{lemma}
\begin{proof}
Suppose that $P\lessdet_T Q$, so that $\trc(Q)\subseteq\trc(P)$, and assume that
$s\in\trc(Q)$ with $\#s < pq$. Since $s$ is also a trace of $P$, we can conlude
$$U_T(pq-1)/s = U_T(pq-1,\#s,G(P)/s)$$ because of (\ref{eq:ifpaT}). Now $\trc(Q)\subseteq\trc(P)$ 
implies $[Q/s]^0\subseteq [P/s]^0 = [G(P)/s]^0$, so $U_T(pq-1,\#s,G(P)/s)$ cannot enter branch (\ref{eq:uta}) and produce a $\efail$-event. Therefore, only four options are
available for the test execution    $(Q\parallel[\Sigma] U_T(p))/s$ to continue.

\medskip
\noindent
{\bf Case~1.} $Q/s$ deadlocks and $\minaccs(G(P)/s) = \{ \varnothing \}$. Then
$U_T(pq-1,\#s,G(P)/s)$ enters branch (\ref{eq:utb}), and the test execution 
stops after event $\epass$.

\medskip
\noindent
{\bf Case~2.} $Q/s$ deadlocks, but $\minaccs(G(P)/s)\neq\{\varnothing\}$. 
Then the whole test execution blocks, and this is means that
neither a $\epass$ nor a $\efail$ event is produced, so the test execution is passed.

\medskip
\noindent
{\bf Case~3.} $Q/s$ selects an event $e\in[Q/s]^0$ and $\#s < pq-1$. Then the test case 
$U_T(pq-1)$ in state
$U_T(pq-1,\#s,G(P)/s)$ can also engage in $e$ by entering branch (\ref{eq:utc}),
and the test execution continues without having produced a $\epass$ or a $\efail$ event.

\medskip
\noindent
{\bf Case~4.} $\#s = pq-1$ holds. Then $U_T(pq-1,\#s,G(P)/s)$ can only enter
branch (\ref{eq:utd}), and the test execution stops after $\epass$.

This case analysis shows that every   execution of $(Q\parallel[\Sigma] U_T(p))$
either stops after $\epass$ or produces neither $\epass$ nor $\efail$. This
proves that $Q$ passes test $U_T(pq-1)$ according to the pass  criterion (\ref{eq:passT}).
\xbox
\end{proof}
% ===============================================================================

\begin{lemma}\label{lemma:mainfexhaustivetrace}
A test suite $\TS_T$ specified as in Theorem~\ref{th:tracetest} is
exhaustive for the fault model specified there and for conformance relation $\lessdet_T$.
\end{lemma}
\begin{proof}
As before in the proofs for
failures testing, we construct the product graph $G=G(P)\times G(Q)$ and recall
that every trace $s\in\trc(P)\cap\trc(Q)$ is associated with a path through $G$ labelled
with the same events as $s$, such that $G/s = (G(P)/s,G(Q)/s)$. Furthermore, we recall from Lemma~\ref{lemma:extendV} that graph state $(G(P)/s,G(Q)/s)$ can always be
reached by a trace $u$ of length less or equal $pq-1$, where  the order of $G(P)$ is $p$ and that of $G(Q)$ is $q$.

Suppose that $P\not\lessdet_T Q$. Then there exists a trace $s\in \trc(Q)\cap\trc(P)$ and an event $e\in [Q/s]^0$ such that $e\not\in [P/s]^0$. Let $u\in \trc(Q)\cap\trc(P)$
be a trace with $\#u < pq$ and $G/u = (G(P)/s,G(Q)/s)$. Then 
$$
U_T(pq-1)/u = U_T(pq-1,\#u,G(P)/s).
$$ 
By assumption, 
$e\in (\Sigma -[P/s]^0) = (\Sigma - [G(P)/s]^0)$. 
Since  $G(Q)/u = G(Q)/s$,
 $Q/u$ can engage into $e$. Then $U_T(pq-1,\#u,G(P)/s)$ will enter branch (\ref{eq:uta}), and the test execution stops after having produced $\efail$. This proves that
  $Q$  fails test $U_T(pq-1)$.
\xbox
\end{proof}