### 2018-11-10 scp-journal/article/main.tex

\documentclass[1p]{elsarticle}

%
%
%\usepackage{lineno,hyperref}
%\modulolinenumbers[5]

\journal{Science of Computer Programming}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
%
\usepackage{url}
\urldef{\mailsa}\path|{peleska,huang}@uni-bremen.de|
\urldef{\mailsb}\path|ana.cavalcanti@york.ac.uk|
\urldef{\mailsc}\path|adenilso@icmc.usp.br|

\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\usepackage{hyperref}

%\usepackage{float}
%\usepackage{lscape}
%\usepackage{lscape}
%
%\usepackage[linesnumbered,lined,boxed]{algorithm2e}
\usepackage[draft]{fixme}
%
\usepackage{zed-csp}
%
%\DeclareMathSymbol{\B}{\mathalpha}{AMSb}{"42}
%\DeclareMathSymbol{\I}{\mathalpha}{AMSb}{"49}
%\DeclareMathSymbol{\N}{\mathalpha}{AMSb}{"4E}
%\DeclareMathSymbol{\Pwr}{\mathalpha}{AMSb}{"50}
%\DeclareMathSymbol{\Q}{\mathalpha}{AMSb}{"51}
%\DeclareMathSymbol{\R}{\mathalpha}{AMSb}{"52}
%\DeclareMathSymbol{\Z}{\mathalpha}{AMSb}{"5A}
%\DeclareMathSymbol{\Sol}{\mathalpha}{AMSb}{"53}
%
%\newcommand{\obsv}[1]{{\cal O}(#1)}
%\newcommand{\ctrl}[1]{{\cal C}(#1)}
%\newcommand{\vdot}[1]{\stackrel{.}{#1}}
%
%\newcommand{\HS}{{\mathcal H}}
%
%\newcommand{\Interval}{\I}
%
\newcommand{\ist}{\mbox{{\tt true}}}
\newcommand{\isf}{\mbox{{\tt false}}}
%\newcommand{\emptytrace}{\langle~\rangle}
%\newcommand{\abegin}{\mathbf{begin}}
%\newcommand{\aend}{\mathbf{end}}
%\newcommand{\alet}{\mathbf{let}}
%\newcommand{\aendlet}{\mathbf{endlet}}
%\newcommand{\ain}{\mathbf{in}}
%\newcommand{\afor}{\mathbf{for}}
%\newcommand{\adownto}{\mathbf{downto}}
%\newcommand{\aforall}{\mathbf{foreach}}
%\newcommand{\awhile}{\mathbf{while}}
%\newcommand{\ado}{\mathbf{do}}
%\newcommand{\aod}{\mathbf{od}}
%\newcommand{\aenddo}{\mathbf{enddo}}
%\newcommand{\aif}{\mathbf{if}}
%\newcommand{\afi}{\mathbf{fi}}
%\newcommand{\athen}{\mathbf{then}}
%\newcommand{\aelse}{\mathbf{else}}
%\newcommand{\aelseif}{\mathbf{elseif}}
%\newcommand{\aendif}{\mathbf{endif}}
%\newcommand{\ainout}{\mathbf{inout}}
%\newcommand{\awhere}{\mathbf{where}}
%\newcommand{\areturn}{\mathbf{return}}
%\newcommand{\aout}{\mathbf{out}}
%\newcommand{\aprocedure}{\mathbf{procedure}}
%\newcommand{\afunction}{\mathbf{function}}
%\newcommand{\abreak}{\mathbf{break}}
%\newcommand{\Sup}[1]{\overline{#1}}
%\newcommand{\Inf}[1]{\underline{#1}}
%
\newcommand{\taba}{\hspace*{3mm}}
%\newcommand{\tabb}{\hspace*{6mm}}
\newcommand{\tabc}{\hspace*{9mm}}
\newcommand{\tabd}{\hspace*{12mm}}
%\newcommand{\tabe}{\hspace*{15mm}}
%\newcommand{\tabf}{\hspace*{18mm}}
%
%\newcommand{\gca}[1]{{{#1}^{\vartriangleright}}}
%\newcommand{\gcb}[1]{{{#1}^{\vartriangleleft}}}
%\newcommand{\gclr}{{{\vartriangleleft\atop\longleftarrow}\atop
%                   {\longrightarrow\atop\vartriangleright}}}
%
%\newcommand{\Nat}{{\mathbb N}}
%\newcommand{\Real}{{\mathbb R}}
%
%\newcommand{\trans}{\longrightarrow}
%\newcommand{\transp}{\longrightarrow_{\power}}
%\newcommand{\transl}{\longrightarrow_{L}}
%\newcommand{\transg}{\longrightarrow_{G}}
%\newcommand{\transcfg}[1]{\stackrel{#1}{\longrightarrow}_{CFG}}
%\newcommand{\isdefd}{=_{\mbox{\footnotesize def}}}
%\newcommand{\equivdef}{\equiv_{\mbox{\footnotesize def}}}
%\newcommand{\mitem}{\mbox{\em M-Item}}
%\newcommand{\currt}{\hat{t}}
%\newcommand{\sigmaa}{\sigma_A}
%\newcommand{\strictimplies}{\stackrel{\bullet}{\Rightarrow}}
%\newcommand{\trl}{/\!/}
%\newcommand{\q}{\textbf{q}}
%\newcommand{\eqc}[2]{[#1;#2]}
%\newcommand{\vest}{V_{\text{\sl est}}}
%\newcommand{\vmax}{V_{\text{\sl MRSP}}}
%\newcommand{\wout}{\mathsf{W}}
%\newcommand{\eout}{\mathsf{EB}}
%\newcommand{\areb}{\mathsf{allowRevokeEB}}
%\newcommand{\sbia}{\mathsf{SBAvailable}}
%\newcommand{\sbz}{\mathsf{sb}_0}
%\newcommand{\ticmd}{\mathsf{TICmd}}
%\newcommand{\dmicmd}{\mathsf{DMICmd}}
%
%\newcommand{\sob}{\mathsf{speedOnBoard}}
%\newcommand{\std}{\mathsf{speedToDriver}}
%\newcommand{\pstd}{\mathsf{permittedSpeedToDriver}}
%\newcommand{\csmsw}{\mathsf{csmSwitch}}
%\newcommand{\sbicmd}{\mathsf{sbiCmd}}
%\newcommand{\sbidisplay}{\mathsf{DMIdisplaySBI}}
%
%
%\newcommand{\dvw}{\mathsf{dV}_{\mathsf{warning}}}
%\newcommand{\dvs}{\mathsf{dV}_{\mathsf{sbi}}}
%\newcommand{\dve}{\mathsf{dV}_{\mathsf{ebi}}}
%
%
%\newcommand{\calcw}{\mathsf{dV\_warning(float)}}
%\newcommand{\calcs}{\mathsf{dV\_sbi(float)}}
%\newcommand{\calce}{\mathsf{dV\_ebi(float)}}
%\newcommand{\calcstd}{\mathsf{calc\_speed\_to\_driver()}}
%\newcommand{\calcpstd}{\mathsf{calc\_permitted\_speed\_to\_driver()}}
%\newcommand{\calcsob}{\mathsf{calc\_speed\_onboard()}}
%
\newcommand{\trc}{\text{\it traces}}
\newcommand{\fails}{\text{\it failures}}
%\newcommand{\qtrc}{\text{qtraces}}
%\newcommand{\iotrc}{\text{L}}
%
%
%%\newtheorem{definition}{Definition}
%%\newtheorem{property}{Property}
%%\newtheorem{lemma}{Lemma}
%%\newtheorem{theorem}{Theorem}
%%\newtheorem{corollary}{Corollary}
%%\newtheorem{property}{Property}
%
%%\newtheorem{note}{Note}[section]
%
\newcommand{\xbox}{\unskip\nobreak\hfil\penalty50
      \hskip2em\hbox{}\nobreak\hfil$\Box$%
      \parfillskip=0pt\finalhyphendemerits=0 \par%

      \medskip
      \parfillskip=0pt plus 1fil
}
%
%\newcommand{\dontshow}[1]{}
%\newcommand{\annot}[1]{\textbf{\textcolor{red}{#1}}}
%\newcommand{\modelsf}{\models_{\text{F}}}
%
%\newcommand{\ta}{\mathbf{A}}
%\newcommand{\te}{\mathbf{E}}
%\newcommand{\tx}{\mathbf{X}}
%\newcommand{\tf}{\mathbf{F}}
%\newcommand{\tg}{\mathbf{G}}
%\newcommand{\tu}{\mathbf{U}}
%\newcommand{\tr}{\mathbf{R}}
%\newcommand{\tw}{\mathbf{W}}
%
%\newcommand{\ttu}[1]{\mathbf{U}^{#1}}
%\newcommand{\ttg}[1]{\mathbf{G}^{#1}}
%\newcommand{\ttf}[1]{\mathbf{F}^{#1}}
%
\newcommand{\epass}{{\text{\it pass}}}
\newcommand{\efail}{{\text{\it fail}}}
%
\newcommand{\pass}{\underline{\text{pass}}}
% \newcommand{\tfail}{\underline{\text{fail}}}
%
%\newcommand{\xpass}[1]{{#1}_\text{pass}}
%\newcommand{\xfail}[1]{{#1}_\text{fail}}
%
%\newcommand{\tc}{{\text{TC}}}
\newcommand{\TS}{{\text{TS}}}
\newcommand{\ii}[1]{\underline{#1}}
%\newcommand{\inc}{\,{\rm I}\,}
%
%\newcommand{\scov}{\text{SCOV}}
%\newcommand{\after}{\text{-after-}}
%\newcommand{\reduction}{\preceq}
%
%\newcommand{\ps}{\le_s}
%\newcommand{\pss}[1]{\stackrel{#1}{\ps}}
%
%\newcommand{\pe}{\sim_s}
%\newcommand{\pes}[1]{\stackrel{#1}{\pe}}
\newcommand{\ol}{\overline}
%\newcommand{\diag}{\mathbf{diag}}
%\newcommand{\primem}{\mathbf{prime}}
%
%\newcommand{\ltl}{\text{\tt LTL}}
%
\newcommand{\accs}{\text{\it Acc}}
\newcommand{\minaccs}{\text{\it minAcc}}
%
\newcommand{\refs}{\text{\it Ref}}
\newcommand{\maxrefs}{\text{\it maxRef}}
\newcommand{\minhits}{\text{\it minHit}}
\newcommand{\minsub}{{\it min}_\subseteq}

%\newcommand{\failure}{{\text {\rm failures}}}

\newcommand{\prefs}{\text{\bf pref}}

\newcommand{\card}[1]{{~|\! #1\!|~}}

\newcommand{\pmax}{P_\text{max}}

\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

\newcounter{examplectr}
\newenvironment{example}[1][]
{\refstepcounter{examplectr}

\medskip
\noindent
{\bf Example~\theexamplectr.}}     %%%\label{#1}
{
%\unskip\nobreak\hfil\penalty50
%      \hskip2em\hbox{}\nobreak\hfil$\Box$%
%      \parfillskip=0pt \finalhyphendemerits=0 \par
}

\newenvironment{proof}[1][]
{

\medskip
\noindent
{\bf Proof.\ }}      
{
%\unskip\nobreak\hfil\penalty50
%      \hskip2em\hbox{}\nobreak\hfil$\Box$%
%      \parfillskip=0pt \finalhyphendemerits=0 \par
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% To avoid formulas in $ $ to be broken across lines
\relpenalty=10000
\binoppenalty=10000

% ===========================================================================

\begin{document}

\begin{frontmatter}

\title{Finite Complete Suites for CSP Refinement Testing}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author{Jan Peleska, Wen-ling Huang \fnref{jpwlh}}
\address{Bremen, Germany}
\fntext[jpwlh]{University of Bremen, \mailsa}

\author{Ana Cavalcanti \fnref{ana}}
\address{York, United Kingdom}
\fntext[ana]{University of York, \mailsb}

%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}
%
%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
%\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{support@elsevier.com}
%
%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
\input{abstract}
\end{abstract}

\begin{keyword}
Model-based testing, CSP\sep Trace Refinement\sep Failures Refinement\sep Complete Test Suites\sep
\end{keyword}
\end{frontmatter}

\input{introduction}

\input{preliminaries}

\input{finitecomplete-failures}

\input{casestudies}

\input{finitecomplete-trace}

\input{complexity}

%%%\input{related}

\input{conclusion}

% ===============================================================================
\section*{References}

\bibliography{references,jp}

\end{document}