### 2018-11-14 scp-journal/article/abstract.tex
Complete version - fit for submission to Science of Computer Programming.

In this paper,   new contributions for model-based testing with
Communicating Sequential Processes (CSP) are presented. For a given CSP
process representing the reference model, finite test suites checking the
conformance relations trace refinement and failures refinement are constructed, and
their   completeness (that is, capability to uncover conformity
violations) is proven. The fault domains for which complete failure detection can be guaranteed
are specified by means of normalised transition graphs representing the failures semantics of finite-state CSP processes.
While complete test suites for CSP processes have been
previously investigated by several authors, a sufficient condition for their
finiteness is presented here for the first time. 
Moreover, is shown that the test suites are optimal with respect to two
aspects:~(a)~the maximal length of test traces cannot be further reduced, and (b)~the
nondeterministic behaviour cannot be tested with smaller or fewer sets of
events, without losing the test suite's completeness property.