### 2018-11-16 scp-journal/article/preliminaries.tex

% =========================================================================
\section{Preliminaries}
\label{section:preliminaries}
% =========================================================================

We present CSP~(Section~\ref{section:csp}) and the concept of minimal hitting
sets~(Section~\ref{sec:hit}), which is central to our notion of test
for failures refinement. To study complexity, we also introduce the concept
of Sperner families~(Section~\ref{sec:sperner}).

% =========================================================================
\subsection{CSP, Refinement, and Normalised Transition Graphs}
\label{section:csp}
% =========================================================================

\paragraph{Communicating Sequential Processes (CSP)} This is a process algebra supporting
system development by refinement. Using CSP, we model both systems and their
components using processes. They are characterised by their patterns of
interactions, modelled by synchronous, instantaneous, and atomic events.

Throughout this paper, the alphabet of the processes, that is, the set of
events in scope, is denoted by $\Sigma$ and supposed to be finite.
The FDR tool~\cite{fdr} supports model checking and semantic analyses of
finite-state CSP processes.

A prefixing operator $e \then P$ defines a process that is ready to engage in
the event $e$, pending agreement of its environment to synchronise. After $e$
occurs, the process behaves as defined by $P$. The environment can be other
processes, in parallel, or the environment of a system as a whole.

Two forms of choice support branching behaviour. An external choice $P
\extchoice Q$ between processes $P$ and $Q$ offers to the environment the
initial events of $P$ and $Q$. Once a synchronisation takes place, the
process that has offered the event that has occurred is chosen and the other
is discarded. In an internal choice $P \intchoice Q$, the environment does
not have an opportunity to interfere:~the choice is made by the process.
%
\begin{example}\label{example:CSP}
  We consider the processes $P$, $Q$, and $R$ defined below. $P$ is initially
  ready to engage in the event $a$, and then makes an internal choice to
  behave like either $Q$ or $R$.
  \begin{eqnarray*}
  P & = & a \then (Q\intchoice R)
  \\
  Q & = & a \then P \extchoice c \then P
  \\
  R & = & b \then P \extchoice c \then R
  \end{eqnarray*}
  $Q$, for instance, offers to the environment the choice to engage in $a$
  again or $c$. In both cases, afterwards, we have a recursion back to $P$.
  In $R$, if $b$ is chosen, we also have a recursion back to $P$. If $c$ is
  chosen, the recursion is to $R$.
  \xbox
\end{example}
%
Iterated forms $\Extchoice i: I @ P(i)$ and $\Intchoice i: I @ P(i)$ of the
external and internal choice operators define a choice over a collection of
processes $P(i)$. If the index set $I$ is empty, the external choice is the
process $\Stop$, which deadlocks:~does not engage into any event or
terminate. For an external choice $\Extchoice e: A @ e \then P(e), A\subseteq
\Sigma$ over a set $A$ of events, we use the abbreviation $e:A \then P(e)$.
An iterated internal choice is not defined for an empty index set.

There are several parallelism operators. A widely used form of parallelism $P
\parallel[cs] Q$ defines a process in which the behaviour is characterised by
those of $P$ and $Q$ in parallel, synchronising on the events in the set
$cs$. Other forms of parallelism available in CSP can be defined using this
parallelism operator.

Interactions that are not supposed to be visible to the environment can be
hidden. The operator $P \hide H$ defines a process that behaves as $P$, with
the interactions modelled by events in the set $H$ hidden. Frequently, hiding
is used in conjunction with parallelism:~it is often desirable to make
actions of each process in a network of parallel processes, perhaps used for
coordination of the network, invisible, while events happening at their
interfaces remain observable.

A rich collection of process operators allows us to define networks of
parallel processes in a concise and elegant way, and reason about safety,
liveness, and divergences.  A comprehensive account of the notation is given
in~\cite{Roscoe2010}.

A distinctive feature of CSP is its treatment of refinement~(as opposed to
bisimulation), which is convenient for reasoning about program correctness,
due to its treatment of nondeterminism and divergence.  A variety of semantic
models capture different notions of refinement. The simplest model
characterises a process by its possible \emph{traces}; the set $\trc(P)$
denotes the sequences of (non-hidden) events in which $P$ can engage.  We say
that a process \emph{$P$ is trace-refined by another process $Q$}, written $P
\lessdet_T Q$, if $\trc(Q)\subseteq \trc(P)$.

In fact, in every semantic model, subset containment is used to define refinement. The
model we focus on first is the failures model, which captures both
sequences of interactions and deadlock behaviour. A \emph{failure} of a process $P$
is a pair $(s,X)$ containing a trace $s$ of $P$ and a \emph{refusal}:~a set $X$ of
events in which $P$ may refuse to engage, after having performed the events of
$s$. The failures model of a process $P$ records all its failures in a set
$\fails(P)$.

Semantic definitions specify, for each operator, how the traces or failures
of the resulting process can be calculated from those of each operand. For
example, for internal choice, $\fails(P\intchoice Q) =
\fails(P)\cup\fails(Q)$; see~\cite[p.~210]{Roscoe:1997:TPC:550448} for a
comprehensive list of these definitions covering $\trc(P)$ and $\fails(P)$.

Using the notation $P/s$ to denote the behaviour ot the process $P$ after
having engaged into the events in the trace $s$, the set $\refs(P/s) \defs
\{~X | (s,X) \in \fails(P)~\}$ contains the  refusals of $P$ after $s$.
Refusals are subset-closed~\cite{Hoare:1985:CSP:3921,Roscoe2010}: if $(s,X)$
is a failure of $P$ and $Y\subseteq X$, then $(s,Y)\in\fails(P)$ and
$Y\in\refs(P/s)$ follows.

For divergence-free processes, failures refinement, $P \lessdet_F Q$, is
defined by $\fails(Q)\subseteq\fails(P)$. Since refusals are subset-closed,
$P \lessdet_F Q$ implies $(s,\varnothing)\in\fails(P)$ for all traces
$s\in\trc(Q)$. So, for divergence-free processes, failures refinement implies
trace refinement. Therefore, using the conformance relation $conf$ below
%
\begin{equation}\label{eq:conf}
  Q\ conf\ P \defs \forall s\in \trc(P) \cap \trc(Q): \refs(Q/s)
  \subseteq \refs(P/s),
\end{equation}
%
failures refinement can be expressed by $\lessdet_T$ and $conf$ as proven
in~\cite{DBLP:conf/icfem/CavalcantiG07}.
%
\begin{equation}%\label{thm:fref-tref-conf}
\label{eq:failconf}
(P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)
\end{equation}
%
For finite processes, since refusals are subset-closed, $\refs(P/s)$ can be
constructed from the set of \emph{maximal refusals}.
%
\begin{equation}
\maxrefs(P/s) = \{ R \in\refs(P/s)~|~\forall R'\in \refs(P/s) - \{ R\}: R \not\subseteq R'\}
\end{equation}
%
Conversely, with the maximal refusals $\maxrefs(P/s)$ at hand, we can
reconstruct the refusals in the set $\refs(P/s)$. % as described by
%
\begin{equation}\label{eq:refMaxRef}
\refs(P/s) = \{ R'\in 2^\Sigma~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
\end{equation}
%
Deterministic process states $P/s$ have exactly the one maximal refusal
defined by $\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of
$P/s$, that is, the events that $P/s$ may engage into. The maximal refusals
in combination with the initials of a process express its degree of
nondeterminism as illustrated by the next example.
%
\begin{example}
\label{ex:nondetdegree} $P = (\Stop \intchoice Q)$ has maximal refusals
$\maxrefs(P) = \{ \Sigma \}$, because $\Stop$ refuses to engage in any event,
and this is carried over to $P$ by the internal choice. However, $P$ is
distinguished from $\Stop$ by its initials, which are defined by $[P]^0 = [
\Stop \intchoice Q]^0 = [Q]^0$. So $P$ may engage nondeterministically in any
initial event of $Q$, but also refuse everything, due to internal selection
of $\Stop$. Assuming an alphabet $\Sigma = \{a,b,c,d\}$, the process
%
$$Q = (e:\{a,b\} \then \Stop) \intchoice (e:\{c,d\} \then \Stop)$$
%
has maximal refusals $\maxrefs(Q) = \{ \{c,d\}, \{a,b\} \}$ and initials
$[Q]^0=\Sigma$. In contrast to $P$, nondeterminism is reflected here by
two maximal refusals. \xbox
\end{example}

% =========================================================================
\paragraph{Normalised Transition Graphs for CSP Processes} \label{sec:ntg}

As shown in~\cite{Roscoe:1994:chapter}, the failures semantics of any
finite-state CSP process $P$ can be represented by a \emph{normalised
transition graph} $G(P)$ defined by a tuple
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
partial \emph{transition function} $t$ maps a node $n$ and an event
$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
$t$, then there is a transition, that is, an outgoing edge, from $n$ with
label $e$, leading to node $t(n,e)$. Normalisation of $G(P)$ is reflected by
the fact that $t$ is a function.

The graph construction in~\cite{Roscoe:1994:chapter} implies that all nodes
$n$ in $N$ are reachable by sequences of edges labelled by $e_1\dots e_k$ and
connecting states $\ii n,n_1,\dots,n_{k-1},n$, such that
\[
n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
n= t(n_{k-1},e_k).
\]
%
By construction, $s\in\Sigma^*$ is a trace of $P$, if, and only if, there is
a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
the events in $s$ in the order they appear. In analogy to $\trc(P)$, we use
the notation $\trc(G(P))$ for the set of finite, initialised paths through
$G(P)$, each path represented by its finite sequence of edge labels. We note
that $\trc(P) = \trc(G(P))$. Since $G(P)$ is normalised, there is a unique
node reached by following the events from $s$ one by one, starting in $\ii
n$. Therefore, $G(P)/s$  is also well defined.

By $[n]^0$ we denote the \emph{initials} of $n$:~the set of events occurring
as labels in any outgoing transitions.
$$
[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
$$
The graph construction from~\cite{Roscoe:1994:chapter} guarantees that $[G(P)/s]^0 = [P/s]^0$ for all
traces $s$ of $P$.

The total function $r$ maps each node $n$ to a non-empty set of  (possibly
empty) subsets of $\Sigma$. The graph construction guarantees that
$r(G(P)/s)$ represents the maximal refusals of $P/s$ for all $s\in\trc(P)$.
As a consequence,
\begin{equation}\label{eq:refsAndR}
(s,X)\in\fails(P) \iff s\in\trc(G(P)) \wedge \exists R\in r(G(P)/s): X\subseteq R,
\end{equation}
so $G(P)$ allows us to re-construct the failures semantics of $P$.
%\fixme{alcc: Why not leave the commented out material in the report?}
% Commented out just to shorten the report.
%To see that this approach works only for finite CSP processes, we consider
%the example where $\Sigma$ is infinite. In this case,
%$\maxrefs(\Stop/\langle\rangle)$ is empty, and so we cannot use this set to
%calculate the refusals of $\Stop$, that is, $\refs(\Stop/\langle\rangle)$ as
%defined above. As with refusals, we also use the transition graph-oriented
%notation $\maxrefs(n) \subseteq r(n)$ to denote the maximal refusals
%associated with graph state $n$: if $n$ is the state reached in the
%transition graph by following the edge labels in trace $s$, then $\maxrefs(n)
%= \maxrefs(P/s)$.

% Commented out just to shorten the report.
%Well-formed normalised transition graphs must not refuse an event of the
%initials of a state in {\it every} refusal applicable in this state; more
%formally,
%%
%\begin{equation}
%\label{eq:wellformedg}
%\forall n\in N, e\in\Sigma: (n,e)\in\dom~t \Rightarrow
%\exists R\in r(n): e\not\in R
%\end{equation}
%%
%By construction, normalised transition graphs reflect the \emph{failures
%semantics} of finite-state CSP processes:~the traces $s$ of a process are
%defined by the sequences of transitions associated with paths through its
%graph, starting at $\ii n$. The pairs $(s,R)$ with $s\in\trc(P)$ and $R\in
%r(G(P)/s)$ represent the failures of $P$.

% =========================================================================
\paragraph{Acceptances} \label{sec:accs}

When investigating tests for failures refinement, the notion of
\emph{acceptances}, which is dual to refusals, is useful. While the
original introduction of acceptances presented
in~\cite[pp.~75]{Hennessy:1988:ATP:50497} was independent of refusals, we use
the definition from~\cite[pp.~278]{Roscoe:1997:TPC:550448}.   A \emph{minimal
acceptance} of a CSP process state $P/s$ is the complement of a maximal
refusal of the same state. The set of minimal acceptances of $P/s$ is denoted
by $\minaccs(P/s)$ and formally defined as
%
\begin{equation}\label{eq:accref}
\minaccs(P/s) = \{ \Sigma - R~|~R\in \maxrefs(P/s)  \}
\end{equation}
%
With this definition, a (not necessarily minimal) \emph{acceptance} of $P/s$ is a superset of some minimal acceptance and a subset of the initials $[P/s]^0$. Denoting the acceptances of $P/s$ by $\accs(P/s)$, this leads to the formal definition
\begin{equation}
\label{eq:accrefall}
\accs(P/s)  = \{ B\subseteq [P/s]^0~|~\exists A\in\minaccs(P/s):  A \subseteq B   \}
\end{equation}
%
Acceptances have the following intuitive interpretation.  If the behaviour of
$P/s$ is deterministic, its only acceptance equals $[P/s]^0$, because $P/s$
never refuses any of the events in this set. If $P/s$ is nondeterministic, it
internally chooses one of its \emph{minimal acceptance sets} $A$ and never
refuses any event in $A$, while {\it possibly} refusing the events from the
set $[P/s]^0 - A$ and {\it always} refusing those in the set $\Sigma -
[P/s]^0$.

Exploiting (\ref{eq:accref}), the nodes of a normalised transition graph can
alternatively be labelled with minimal acceptances; this captured the same
information conveyed by maximal refusals. Since process states $P/s$ are
equivalently expressed by states $G(P)/s$ of $P$'s normalised transition
graph, we also write $\minaccs(G(P)/s)$ and note that (\ref{eq:refsAndR}) and
(\ref{eq:accref}) imply
\begin{equation}\label{eq:maxrefsminaccs}
\minaccs(G(P)/s) = \{ \Sigma - R | R\in r(G(P)/s) \} = \minaccs(P/s).
\end{equation}
%
Given any non-diverging, non-terminating, finite-state process $P$, it can be
re-con\-struc\-ted from its graph $G(P)$ with initial state $\ii
n$ and transition function $t$, using $P$'s normalised syntactic
representation~\cite[pp.~277]{Roscoe:1997:TPC:550448} specified as follows.
\begin{eqnarray*}
\text{normalised}(P) & = & P_N(\ii n)
\\
P_N(n) & = & \Intchoice_{A\in\minaccs(n)\cup \{ [n]^0 \} } e:A\then P_N(t(n,e))
\end{eqnarray*}
With this definition, it is established that $P$ is semantically equivalent
to $\text{normalised}(P)$ in the failures semantics.

% .....................................................................................
 \begin{figure}[t]
   %%\hspace*{-40mm}
   \begin{center}
     \includegraphics[trim=0cm 1cm 0cm 1cm,clip,scale=.6,%width=\textwidth
                              ]{p.pdf}
   \end{center}
   %%\vspace*{-10mm}
   \caption{Normalised transition graph of CSP process $P$ from Example~\ref{ex:a}.}
   \label{fig:tga}
 \end{figure}
% .......................................................................................

\begin{example}\label{ex:a}
We consider the process $P$ in Example~\ref{example:CSP}; its transition
graph $G(P)$ is shown in Fig.~\ref{fig:tga}. The process state
$P/\varepsilon$ (where $\varepsilon$ denotes the empty trace) is represented
as node 0, with $\{ a\}$ as the only minimal acceptance, since $a$ is never
refused and no other events are accepted. Having engaged in $a$, the
transition from node 0 leads to node 1 representing the process state $P/a
= Q\intchoice R$. The internal choice induces several minimal acceptances
derived from $Q$ and $R$. Since these processes accept their initial events
in external choice, $Q\intchoice R$ induces minimal acceptance sets $\{a,c\}$
and $\{b,c\}$. We note that the event $c$ can never be refused, since it is
contained in each minimal acceptance set.

Having engaged in $c$, the next process state is represented by node 2. Due
to normalisation, there is only a single transition satisfying $t(1,c) = 2$.
This transition, however, can have been caused by either $Q$ or $R$ engaging
into $c$, so node 2 corresponds to process state $Q/c \intchoice R/c = P
\intchoice R$. This is reflected by the two minimal acceptance sets labelling
node 2. From node~2, event $c$ leads to node~3. Since $P$ does not engage
into $c$, the $R$-component of $P\intchoice R$ must have processed $c$, so
node~3 corresponds to $R/c = R$, and so it is labelled by $R$'s minimal
acceptance $\{b,c\}$. \xbox
\end{example}
%
\noindent%
Summarising, refinement between finite-state CSP processes $P, Q$ can be
expressed using their normalised graphs
\begin{eqnarray*}
G(P) & = & ( N_P, {\ii n}_P, \Sigma, t_P : N_P\times\Sigma \pfun N_P, r_P : N_P \fun \mathbb{P}\mathbb{P}(\Sigma))
\\
G(Q) & = & ( N_Q, {\ii n}_Q, \Sigma, t_Q : N_Q\times\Sigma \pfun N_Q, r_Q : N_Q \fun \mathbb{P}\mathbb{P}(\Sigma))
\end{eqnarray*}
%
as established by the results  in the following lemma. There, result
(\ref{eq:trcrefa}) reflects trace refinement in terms of graph traces; result
(\ref{eq:failconf}) expresses failures refinement in terms of traces
refinement and $conf$; result (\ref{eq:failrefa}) states how $conf$ can be
expressed by means of the maximal refusal functions of the graphs involved;
and result (\ref{eq:failrefb}) states the same in terms of the minimal
acceptances that can be derived from the maximal refusal functions by means
of (\ref{eq:maxrefsminaccs}).
%
\begin{lemma}
  \label{lemma:tgtrcref}
  \begin{eqnarray}
  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
  \label{eq:trcrefa}
%  \\
%  \label{eq:failconf}
%  P \lessdet_F Q & \Leftrightarrow & P \lessdet_T Q \wedge Q\ conf\ P
  \\
  \label{eq:failrefa}
  Q\ conf\ P & \Leftrightarrow & \nonumber
  \forall s\in\trc(G(Q))\cap \trc(G(P)), R_Q\in r_Q(G(Q)/s):  \nonumber
  \\ & & \tabd
  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
  \\
  \label{eq:failrefb}
   & \Leftrightarrow &
    \forall s\in\trc(G(Q))\cap \trc(G(P)), A_Q\in\minaccs(G(Q)/s):  \nonumber
   \\ & & \tabd
  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
 \end{eqnarray}
  \xbox
\end{lemma}
%
\begin{proof}
To prove (\ref{eq:trcrefa}), we recall that $P\lessdet_T Q$ is defined as
$\trc(Q)\subseteq \trc(P)$ and, moreover, $\trc(P)=\trc(G(P))$ and
$\trc(Q)=\trc(G(Q))$. %Result (\ref{eq:failconf}) has been shown
%in~\cite{DBLP:conf/icfem/CavalcantiG07}.
%
%\medskip
%\noindent
%To prove (\ref{eq:failconf}), we derive
%\begin{argue}
%P \lessdet_F Q
%\\
%\Leftrightarrow \fails(Q)\subseteq \fails(Q) & definition of $\lessdet_F$
%\\
%\Leftrightarrow \forall (s,R) \in  \fails(Q): (s,R) \in\fails(P)
%& property of $\subseteq$
%\\
%\Leftrightarrow \forall (s,\varnothing) \in  \fails(Q): (s,\varnothing) \in\fails(P)
%\wedge {}
%\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
%\\
%& Definition of failures, refusals, subset closure
%\\
%\Leftrightarrow \trc(Q)\subseteq\trc(P)
%\wedge {}
%\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
%\\
%& Definition of failures and traces
%\\
%\Leftrightarrow \trc(Q)\subseteq\trc(P)
%\wedge {}
%\\\tabc \forall s\in\trc(Q)\cap\trc(P): \refs(Q/s) \subseteq\refs(P/s)
%\\
%& Property of $\cap$ and $\subseteq$
%\\
%\Leftrightarrow
%P\lessdet_T Q \wedge Q\ conf\ P
%& Definition of $\lessdet_T$ and $conf$
%\end{argue}
%
To prove (\ref{eq:failrefa}), we derive
\begin{argue}
  Q\ conf\ P
  \\
  \Leftrightarrow \forall s\in\trc(P)\cap \trc(Q): \refs(Q/s)\subseteq\refs(P/s)
  %\\
  & Definition of $conf$  (\ref{eq:conf})
  \\
  \Leftrightarrow \forall s\in\trc(G(P))\cap \trc(G(Q)): \refs(Q/s)\subseteq\refs(P/s)
  \\
  & $\trc(P) = \trc(G(P))$, $\trc(Q) = \trc(G(Q))$
  \\
  \Leftrightarrow \forall s\in\trc(G(P))\cap \trc(G(Q)), R_Q\in r_Q(G(Q)/s):
  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
  \\ %
  & Property of $r_P, r_Q$ and (\ref{eq:refMaxRef})
\end{argue}

\medskip
\noindent
Finally, (\ref{eq:failrefb}) follows from (\ref{eq:failrefa}) using (\ref{eq:accref})
and the fact that $R_Q \subseteq R_P$ is equivalent to
$\Sigma - R_P \subseteq \Sigma - R_Q$.
\xbox
\end{proof}

% =========================================================================
\paragraph{Reachability Under Sets of Traces} \label{sec:V} Given a finite-state CSP
process $P$ and its normalised transition graph $G(P)$,
%\[
%G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
%\]
we suppose that $V\subseteq\Sigma^*$ is a prefix-closed set  of sequences of
events. By $t(\ii n,V)$ we denote the set
\[
t(\ii n,V) = \{ n\in N~|~\exists s\in V: s\in\trc(P)\wedge G(P)/s = n \}
\]
of nodes in $N$ that are reachable in $G(P)$ by applying traces of $V$. The
lemma below specifies a construction method for such sets $V$ reaching {\it
every} node of $N$.

\begin{lemma}
\label{lemma:extendV} Let $P$ be a CSP process with normalised transition
graph $G(P)$. Let $V\subseteq\Sigma^*$ be a finite prefix-closed set of
sequences of events. Suppose that  $G(P)$ reaches $k < |N|$ nodes under $V$,
that is, $|t(\ii n,V)| = k$. Let $V.\Sigma$ denote the set of all sequences
from $V$, extended by any event of $\Sigma$.  Then $G(P)$ reaches at least
$(k+1)$ nodes under $V\cup V.\Sigma$.
\end{lemma}
\begin{proof}
Suppose that $n'\in (N - t(\ii n,V))$.  Since all nodes in $N$ are reachable,
there exists a trace $s$ such that $G(P)/s = n'$. Decompose $s = s_1.e.s_2$
with $s_i\in\Sigma^*, e\in\Sigma$, such that $G(P)/s_1 \in t(\ii n,V)$ and
$G(P)/s_1.e \not\in t(\ii n,V)$. Such a decomposition always exists, because
$V$ is prefix-closed and therefore contains the empty trace $\varepsilon$.
Note, however, that it is not necessarily the case that $s_1\in V$. Since
$G(P)$ reaches $G(P)/s_1$ under $V$, there exists a trace $u\in V$ such that
$G(P)/u = G(P)/s_1 = \ol n$. Since $s = s_1.e.s_2$ is a trace of $P$ and
$G(P)/s_1 = \ol n$, then $(\ol n,e)$ is in the domain of $t$. So, $ G(P)/u.e
= G(P)/s_1.e = n$ is a well-defined node of $N$ not contained in $t(\ii
n,V)$. Since $u.e\in V\cup V.\Sigma$, $G(P)$ reaches at least the additional
node $n$ under $V\cup V.\Sigma$. This completes the proof. \xbox
\end{proof}

% =========================================================================
\paragraph{Graph Products} \label{sec:GP} For proving our main theorems, it is
necessary to consider the \emph{product} of normalised transition graphs. We
need this only for the investigation of corresponding traces in reference
processes and processes for SUTs. So, the labelling of nodes with maximal
refusals or minimal acceptances are disregarded in the product construction.
We consider two normalised transition graphs
\[
G_i = ( N_i, \ii n_i, \Sigma, t_i : N_i\times\Sigma \pfun N_i, r_i : N_i \fun \mathbb{P}\mathbb{P}(\Sigma)),\qquad i = 1,2,
\]
over the same alphabet $\Sigma$. Their product is defined by
%
\begin{eqnarray}
G_1\times G_2 & = & (N_1\times N_2,(\ii n_1,\ii n_2), t:(N_1\times N_2)\times\Sigma\pfun (N_1\times N_2))
\\
\dom~t & = & \{ ((n_1,n_2),e)\in (N_1\times N_2)\times\Sigma | (n_1,e)\in\dom~t_1\wedge (n_2,e) \in\dom~t_2 \}
\\
t((n_1,n_2),e) & = & (t_1(n_1,e),t_2(n_2,e))\ \text{for $((n_1,n_2),e)\in\dom~t$}
\end{eqnarray}
%
The following lemma is used in the proof of our main theorem.
%
\begin{lemma}\label{lemma:reachproduc}
If $G_1$ has $p$ states and $G_2$ has $q$ states, then every reachable state
$(n_1,n_2)$ of the product graph $G_1\times G_2$ can be reached by a trace
%%%$s\in\trc(G_1)\cap \trc(G_2)$  we don't need this
of maximal length $(pq-1)$.
\end{lemma}
\begin{proof}
The product graph $G_1\times G_2$ has at most $pq$ states. The empty trace $\varepsilon$
reaches its initial state $(\ii n_1,\ii n_2)$. Applying Lemma~\ref{lemma:extendV}
$(pq-1)$ times with $V=\{\varepsilon \}$ implies that $G_1\times G_2$ reaches
all of its reachable states (there are at most $pq$ of them) under
$V' = V \cup V.\Sigma\cup\dots \cup V.\Sigma^{(pq-1)}$. The maximal length of traces in
$V'$ is $(pq-1)$.
\xbox
\end{proof}
%
This concludes our presentation of CSP and of results regarding its
semantics that are used in the next section.

% =========================================================================
\subsection{Minimal Hitting Sets}
\label{sec:hit}
% =========================================================================

\paragraph{Definition} The main idea of the underlying test strategy for failures
refinement is based on solving a \emph{hitting set problem}. Given a finite
collection of finite sets ${\cal C} = \{ A_1,\dots,A_n\}$, such that each
$A_i$ is a subset of a universe $\Sigma$, a \emph{hitting set}
$H\subseteq\Sigma$ is a set satisfying the following property.
%
\begin{equation}
  \label{eq:hit}
  \forall A\in {\cal C}: H\cap A \neq\varnothing.
\end{equation}
%
A \emph{minimal hitting set} is a hitting set that cannot be further reduced
without losing the characteristic property (\ref{eq:hit}). By $\minhits({\cal
C})$ we denote the collection of minimal hitting sets for a collection ${\cal
C}$. For the pathological case where $C$ contains an empty set,
$\minhits({\cal C})$ is also empty. The problem of determining minimal hitting sets is %%% \cite{Book1975-BOOKRM,5533149}
NP-hard~\cite{5533149}. We see below, however, that using minimal hitting
sets, we can reduce the effort of testing for failures refinement from a
factor of $2^{|\Sigma|}$ to a factor that equals the number of minimal
hitting sets.

% -----------------------------------------------------------------------------
\paragraph{Minimal Hitting Sets of Minimal Acceptances} In this article, we are
interested in the minimal hitting sets of minimal acceptances; for these, the
abbreviated notation
$%\[ \
minhits(P/s) = \minhits(\minaccs(P/s))
$ %\]
is used. The minimal hitting sets
of minimal acceptances may be alternatively characterised by means of the
failures of a process as is done in~\cite{DBLP:conf/icfem/CavalcantiG07}. To
this end, in~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors define, for
any collection ${\cal C}\subseteq 2^\Sigma$ of subsets from $\Sigma$
\begin{equation}\label{eq:minsub}
\minsub({\cal C}) = \{ A\in {\cal C}~|~\forall B\in {\cal C}:
B\subseteq A\Rightarrow B = A \}.
\end{equation}
The collection $\minsub({\cal C})$ contains all those sets of
${\cal C}$ that are not true subsets of other members of ${\cal C}$. With
this definition, the relation between failures and minimal hitting sets of
minimal acceptances is established in the following lemma.
%
\begin{lemma}
\label{lemma:failureshittingsets}
For any trace $s$ of CSP process $P$, define
$%\[
{\cal A}_s = \{ A\subseteq \Sigma~|~(s,A)\not\in\fails(P)  \}.
$ %\]
Then
$ %\[
\minhits(P/s) = \minsub {\cal A}_s
$ %\]
for all traces $s$ of $P$.
\end{lemma}
\begin{proof}
We derive
\begin{argue}
(s,A)\not\in {\text{failures}}(P)
\\
\Leftrightarrow A\not\in {\text{Ref}}(P/s)
\\
\Leftrightarrow  \forall R\in {\text{maxRef}}(P/s): A\not\subseteq R
\\
\Leftrightarrow   \forall B\in {\text{minAcc}}(P/s): A\not\subseteq (\Sigma\setminus B)& this follows from (\ref{eq:accref})
\\
\Leftrightarrow  \forall B\in {\text{minAcc}}(P/s): A\cap B\neq \varnothing
\\
\Leftrightarrow  A \text{ is a (not necessarily minimal)}
              \text{ hitting set of } {\text{minAcc}}(P/s)
\end{argue}
%
Specialising this derivation valid for arbitrary $A$ satisfying $(s,A)\not\in {\text{failures}}(P)$ to
minimal $A$ with this property proves the statement of the lemma.
\xbox
\end{proof}

% -------------------------------------------------------------------------
\paragraph{Minimal Hitting Sets of Normalised Transition Graphs} Since, as
previously explained, minimal acceptances can be used to label the nodes of a
normalised transition graph, and since $\minaccs(P/s) = \minaccs(G(P)/s)$ by
(\ref{eq:maxrefsminaccs}), the notation of minimal hitting sets also carries
over to graphs: we write $\minhits(n)$ for nodes $n$ of $G(P)$ and observe
that
\begin{equation}
\label{eq:minhitsGP}
\minhits(G(P)/s) = \minhits(P/s) \ \text{for all}\ s\in \trc(P).
\end{equation}

% ------------------------------------------------------------------------
\paragraph{Characterisation of $conf$ by Minimal Hitting Sets} The following lemma
establishes that the $conf$ relation specified in (\ref{eq:conf}) can be
characterised by means of minimal acceptances and their minimal hitting sets.
%
\begin{lemma}
\label{lemma:hseta}
Let $P, Q$ be two finite-state CSP processes.
%% satisfying $P\lessdet_T Q$.
For each $s\in\trc(P)$,
let $\minhits(P/s)$ denote the
collection of all minimal hitting sets of $\minaccs(P/s)$.
Then the following statements are equivalent.
\begin{enumerate}
\item $Q\ conf\ P$

\item For all $s\in\trc(P)\cap \trc(Q)$ and $H \in  \minhits(P/s)$, $H$ is
a (not necessariliy minimal) hitting set of $\minaccs(Q/s)$.
\end{enumerate}
\end{lemma}
\begin{proof}
We apply (\ref{eq:maxrefsminaccs}) and (\ref{eq:minhitsGP}), so that
$\minaccs(P/s)$ and $\minaccs(G(P)/s)$, as well as $\minhits(P/s)$ and
$\minhits(G(P)/s)$ are used interchangeably. For showing ``$(1) \Rightarrow
(2)$'', we assume $Q\ conf\ P$ and $s\in\trc(P)\cap \trc(Q)$.
Lemma~\ref{lemma:tgtrcref} (\ref{eq:failrefb}), states that
$%\[
\forall A_Q\in\minaccs(G(Q)/s):
\exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
$. %\]
Therefore, $H \in  \minhits(P/s)$ not only implies $H\cap A_P\neq\varnothing$
for all minimal acceptances $A_P$, but also $H\cap A_Q\neq\varnothing$ for
every minimal acceptance $A_Q$, because $A_P\subseteq A_Q$ for at least one
$A_P$. So, each $H \in \minhits(P/s)$ is also a hitting set for
$\minaccs(G(Q)/s)$ as required.

To prove ``$(2) \Rightarrow (1)$'', we assume that (2) holds, but that $P\
conf\ Q$ does {\it not} hold. According to Lemma~\ref{lemma:tgtrcref},
(\ref{eq:failrefb}), there exists $s\in\trc(P)\cap \trc(Q)$ such that
\[
\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s):
A_P\not\subseteq A_Q \qquad (*)
\]
Let $A$ be such an acceptance set $A_Q$ fulfilling (*).
Define
$%\[
\overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
$. %\]
Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),
$\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
intersection with $A$. Minimising $\overline H$ yields   a minimal hitting
set $H\in \minhits(P/s)$ which is {\it not} a hitting set of
$\minaccs(G(Q)/s)$, a contradiction to Assumption~2.
This completes the proof of the lemma.
\xbox
\end{proof}
%
We note that $\minaccs(P) = \{ \varnothing \}$ if $P = Q\intchoice \Stop$.
Since $\Stop$ accepts nothing, its minimal acceptance is $\varnothing$, and
this carries over to $Q\intchoice \Stop$.  From (\ref{eq:failrefb}) we
conclude that $\varnothing\in\minaccs(P)$ implies $\minaccs(P) = \{
\varnothing \}$. This clarifies that $\minhits(P/s)$ is empty if, and only
if, $\minaccs(P) = \{ \varnothing \}$. The proof of Lemma~\ref{lemma:hseta}
covers the situations where $\minaccs(P/s) = \{ \varnothing \}$ and so
$\minhits(P/s) = \varnothing$. Trivially,
\begin{equation}
\label{eq:minhitminaccempty}
\minaccs(P/s) = \{ \varnothing \} \iff \minhits(P/s) = \varnothing
\end{equation}
holds.

% =========================================================================
\subsection{Sperner Families}
\label{sec:sperner}
% =========================================================================

In preparation for complexity results presented in
Section~\ref{sec:complexity}, we consider how many minimal hitting sets can
maximally exist for a collection of minimal acceptances. To this end, the
following definitions and results are useful.

A \emph{Sperner Family} is a collection ${\cal S} \subseteq 2^\Sigma$ of sets
from a given finite universe $\Sigma$ that do not contain each other, that
is, $H_1 \not\subseteq H_2 \wedge H_2 \not\subseteq H_1$ holds for each pair
$H_1\neq H_2\in {\cal S}$. Specialising antichains known from partial orders
to finite sets partially ordered by $\subseteq$ results in Sperner families.
Given an {\it arbitrary} collection of subsets ${\cal C}\subset 2^\Sigma$,
the sub-collection $\minsub({\cal C})$ defined in (\ref{eq:minsub}) is a
Sperner Family contained in ${\cal C}$.

\newpage
\noindent
We further observe that
\begin{itemize}
\item the maximal refusals of a CSP process state,
\item the minimal acceptances of a CSP process state, and
\item the minimal hitting sets of a given collection of sets
\end{itemize}
are Sperner families. Moreover, given any finite alphabet $\Sigma$ with
$\card{\Sigma} = n$, every collection $S$ of subsets with identical
cardinality $k \le n$ is a Sperner family,  because $A_1, A_2\in S \wedge A_1
\subseteq A_2\wedge \card{A_1} = \card{A_2}$ implies $A_1 = A_2$. Given  any
Sperner Family $S$ of $\Sigma$, $S$ represents the minimal acceptances in the
initial state of the CSP process
%\[
$P = \Intchoice_{A\in S}\big( e:A \then P(e) \big)$.
%\]

The cardinality of Sperner Families is determined by the following theorem.
\begin{theorem}[Sperner's Theorem~\cite{sperner_satz_1928}]
\label{th:sperner} Given a  Sperner family ${\cal S}$ over an $n$-element
universe $\Sigma$, its cardinality $\card{{\cal S}}$ is bound by
\[
\card{{\cal S}} \le \binom{n}{\lfloor\frac{n}{2}\rfloor}.
\]
The upper bound is reached if, and only if, one of the following cases apply:
\begin{enumerate}
\item For even $n$, if ${\cal S}$ consists of all subsets of $\Sigma$ with
    cardinality $n/2$;
\item For odd $n$, if one of the following cases holds;
\begin{enumerate}
\item ${\cal S}$ consists of all subsets of $\Sigma$ with cardinality $(n+1)/2$;
    or
\item ${\cal S}$ consists of all subsets of $\Sigma$ with cardinality $(n-1)/2$.
\end{enumerate}
\end{enumerate}
\xbox
\end{theorem}
It is shown in Section~\ref{sec:complexity} that this upper bound can
actually be reached by the Sperner Family containing the hitting sets
associated with the minimal acceptances of a CSP process state.