### 2018-11-30 scp-journal/article-before-submission/abstract.tex
All files as completed BEFORE submission

In this paper, new contributions for model-based testing using Communicating
Sequential Processes (CSP) are presented. For a finite non-terminating CSP
process representing the reference model, finite test suites for checking the
conformance relations traces and failures refinement are presented, and their
completeness (that is, capability to uncover conformity violations) is
proven. The fault domains for which complete failure detection can be
guaranteed are specified by means of normalised transition graphs
representing the failures semantics of finite-state CSP processes. While
complete test suites for CSP processes have been previously investigated by
several authors, a sufficient condition for their finiteness is presented
here for the first time. Moreover, it is shown that the test suites are
optimal in two aspects:~(a)~the maximal length of test traces cannot be
further reduced, and (b)~the nondeterministic behaviour cannot be tested with
smaller or fewer sets of events, without losing the test suite's completeness
property.